---
title: "Data Modelling V42 (New time and wind variables + weight for XGB (removed 2021-04-27 2 GStreet))"
author: "Jerry Wu"
date: "2026-01-12"
output:
  html_document:
    df_print: paged
  pdf_document: default
abstract: ''
subtitle: ''
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(caret)
library(xgboost)
library(circular)
library(raster)
library(terra)
library(sf)
library(ggpubr)
library(ggpmisc)
library(leaps)
library(kableExtra)
library(gginnards)
library(patchwork)
library(arrow)
select <- dplyr::select
```

# Load in data {.tabset .tabset-pills}
## Full data
```{r}
hourly_full <- readRDS('../../data/hourly.rds')
daily_from_hour_avg_full <- readRDS('../../data/daily_from_hourly_avg.rds')
```

## Splitted data
```{r}
daily_ha_train <- readRDS('../../data/daily_ha_train.rds')
daily_ha_test <- readRDS('../../data/daily_ha_test.rds')
hourly_train <- readRDS('../../data/hourly_train.rds')
hourly_test <- readRDS('../../data/hourly_test.rds')
```

```{r}
# Rename columns for clarity, as it's already transformed.
rename_dist <- c('inv_dist_ref_p2' = 'dist_ref',
                 'inv_dist_dc_p2' = 'dist_dc',
                 'inv_dist_wrp_p2' = 'dist_wrp')

daily_ha_train <- daily_ha_train %>%
  rename(rename_dist)
daily_ha_test <- daily_ha_test %>%
  rename(rename_dist)
hourly_train <- hourly_train %>%
  rename(rename_dist)
hourly_test <- hourly_test %>%
  rename(rename_dist)
```

## Add new variables
```{r}
deg2rad <- function(x) x * pi / 180

make_features <- function(df,
                          id_col = "Monitor",
                          time_col = "timestamp",
                          wd_col = 'wd_avg', 
                          ws_col = 'ws_avg') {
   
    df2 <- df %>%
      group_by(.data[[id_col]]) %>%
      arrange(.data[[time_col]], .by_group = TRUE) %>%
      mutate(
        # time parts
        hour   = hour(.data[[time_col]]),          # 0 to 23
        month  = month(.data[[time_col]]),         # 1 to 12
        julian = yday(.data[[time_col]]),          # 1 to 365/366
  
        # cyclic time (fourier)
        hour_sin   = sin(2 * pi * hour / 24),
        hour_cos   = cos(2 * pi * hour / 24),
        weekday_sin = sin(2 * pi * weekday / 7),
        weekday_cos = cos(2 * pi * weekday / 7),
        month_sin  = sin(2 * pi * month / 12),
        month_cos  = cos(2 * pi * month / 12),
        julian_sin = sin(2 * pi * julian / 365.25),
        julian_cos = cos(2 * pi * julian / 365.25),
  
        # wind transforms
        wd_rad = deg2rad(.data[[wd_col]]),
        wd_sin = sin(wd_rad),
        wd_cos = cos(wd_rad)#,
        # wind_u = .data[[wd_col]] * wd_cos,
        # wind_v = .data[[ws_col]] * wd_sin
      ) %>%
      ungroup()

  
  df2 %>%
      select(-wd_rad)  
}

# run them on training and test
hourly_train <- hourly_train %>%
  mutate(timestamp = as.POSIXct(dayhour, format = "%Y-%m-%d %H", tz = "America/Los_Angeles"))

hourly_test <- hourly_test %>%
  mutate(timestamp = as.POSIXct(dayhour, format = "%Y-%m-%d %H", tz = "America/Los_Angeles"))

hourly_train <- make_features(hourly_train, id_col = "Monitor", time_col = "timestamp")
hourly_test <- make_features(hourly_test, id_col = "Monitor", time_col = "timestamp")

# Compute for daily average features 
daily_ha_train <- make_features(daily_ha_train, id_col = "Monitor", time_col = "day", 
                                wd_col = 'daily_wd', ws_col = 'daily_ws')
daily_ha_test <- make_features(daily_ha_test, id_col = "Monitor", time_col = "day", 
                               wd_col = 'daily_wd', ws_col = 'daily_ws')

daily_ha_train <- daily_ha_train %>%
  rename('daily_wd_sin' = 'wd_sin',
         'daily_wd_cos' = 'wd_cos')

daily_ha_test <- daily_ha_test %>%
  rename('daily_wd_sin' = 'wd_sin',
         'daily_wd_cos' = 'wd_cos')

# Compute for daily max features
daily_ha_train <- make_features(daily_ha_train, id_col = "Monitor", time_col = "day", 
                                wd_col = 'hourly_wd', ws_col = 'hourly_ws')
daily_ha_test <- make_features(daily_ha_test, id_col = "Monitor", time_col = "day", 
                               wd_col = 'hourly_wd', ws_col = 'hourly_ws')

daily_ha_train <- daily_ha_train %>%
  rename('hourly_wd_sin' = 'wd_sin',
         'hourly_wd_cos' = 'wd_cos')

daily_ha_test <- daily_ha_test %>%
  rename('hourly_wd_sin' = 'wd_sin',
         'hourly_wd_cos' = 'wd_cos')
```

```{r}
# Exclude Event stepwise
hourly_excl_dis_train <- hourly_train %>% 
  filter(!(year == '2021' & month %in% c('10', '11', '12')))
daily_ha_excl_dis_train <- daily_ha_train %>% 
  filter(!(year == '2021' & month %in% c('10', '11', '12')))

hourly_excl_dis_test<- hourly_test %>% 
  filter(!(year == '2021' & month %in% c('10', '11', '12')))
daily_ha_excl_dis_test <- daily_ha_test %>% 
  filter(!(year == '2021' & month %in% c('10', '11', '12')))

# Everything w.o disaster indicator
hourly_full_train <- hourly_train %>%
  select(-disaster)
daily_ha_full_train <- daily_ha_train %>%
  select(-disaster)

hourly_full_test <- hourly_test %>%
  select(-disaster)
daily_ha_full_test <- daily_ha_test %>%
  select(-disaster)
```


# XGBoost {.tabset .tabset-pills}
## Helpers and setup {.tabset .tabset-pills}
```{r}
# Save path
model_folder <- '../../models/XGB/' # For all models
```

```{r}
# Result tables
test_result <- tibble(
  Model = character(),
  model_response_name = character(),
  daterange = character(),
  "R2" = numeric(),
  "RMSE" = numeric(),
  "coef" = numeric(),
  "BT_R2" = numeric(),
  "BT_RMSE" = numeric(),
  "BT_coef" = numeric(),
  "event_R2" = numeric(),
  "event_RMSE" = numeric(),
  "event_coef" = numeric(),
  "normal_R2" = numeric(),
  "normal_RMSE" = numeric(),
  "normal_coef" = numeric(),
  "BT_event_R2" = numeric(),
  "BT_event_RMSE" = numeric(),
  "BT_event_coef" = numeric(),
  "BT_normal_R2" = numeric(),
  "BT_normal_RMSE" = numeric(),
  "BT_normal_coef" = numeric()
)

spatial_test_result <- tibble(
  Model = character(),
  model_response_name = character(),
  daterange = character(),
  "lag" = character(),
  "coordinates" = character(),
  "R2" = numeric(),
  "RMSE" = numeric(),
  "coef" = numeric(),
  "R2_mean" = numeric(),
  "RMSE_mean" = numeric(),
  "coef_mean" = numeric(),
  "event_R2" = numeric(),
  "event_RMSE" = numeric(),
  "event_coef" = numeric(),
  "normal_R2" = numeric(),
  "normal_RMSE" = numeric(),
  "normal_coef" = numeric()
)

spatial_weighted_test_result <- tibble(
  Model = character(),
  model_response_name = character(),
  daterange = character(),
  "lag" = character(),
  "coordinates" = character(),
  "R2" = numeric(),
  "RMSE" = numeric(),
  "coef" = numeric(),
  "R2_mean" = numeric(),
  "RMSE_mean" = numeric(),
  "coef_mean" = numeric(),
  "event_R2" = numeric(),
  "event_RMSE" = numeric(),
  "event_coef" = numeric(),
  "normal_R2" = numeric(),
  "normal_RMSE" = numeric(),
  "normal_coef" = numeric()
)

temporal_test_result <- tibble(
  Model = character(),
  model_response_name = character(),
  daterange = character(),
  "R2" = numeric(),
  "RMSE" = numeric(),
  "coef" = numeric()
)

test_result_unstrat <- test_result

negative_pred <- tibble(Model = character(),
  model_response_name = character(),
  daterange = character(),
  "negative" = numeric())
```

```{r}
daha_names <- c(
  "Average Daily Temperature" = "daily_temp",
  "Average Daily Humidity" = "daily_hum",
  "Daily Precipitation" = "daily_precip",
  "Average Daily Wind Speed" = "daily_ws",
  "Average Daily Wind Direction (sin)" = "daily_wd_sin",
  "Average Daily Wind Direction (cos)" = "daily_wd_cos",
  "Downwind Refinery" = "daily_downwind_ref",
  "Downwind WRP" = "daily_downwind_wrp"
)

ha_names <- c(
  "Hourly Temperature" = "hourly_temp",
  "Hourly Humidity" = "hourly_hum",
  "Hourly Precipitation" = "hourly_precip",
  "Hourly Wind Speed" = "ws_avg",
  "Hourly Wind Direction (sin)" = "wd_sin",
  "Hourly Wind Direction (cos)" = "wd_cos",
  "Downwind Refinery" = "hourly_downwind_ref",
  "Downwind WRP" = "hourly_downwind_wrp"
)

dmha_names <- c(
  "Hourly Temperature at peak H2S" = "hourly_temp",
  "Hourly Humidity at peak H2S" = "hourly_hum",
  "Hourly Precipitation at peak H2S" = "hourly_precip",
  "Hourly Wind Speed at peak H2S" = "hourly_ws",
  "Hourly Wind Direction at peak H2S (sin)" = "hourly_wd_sin",
  "Hourly Wind Direction at peak H2S (cos)" = "hourly_wd_cos",
  "Downwind Refinery at peak H2S" = "hourly_downwind_ref",
  "Downwind WRP at peak H2S" = "hourly_downwind_wrp",
  "Hourly Wind Speed"  = "hourly_ws_avg"
)

fixed_names <- c(
  "Longitude" = "mon_utm_x",
  "Latitude" = "mon_utm_y",
  "Distance to Refinery" = "dist_ref",
  "Angle to Refinery" = "angle_ref",
  "Active Wells within 2km" = "active_2km",
  "Inactive Wells within 2km" = "inactive_2km",
  "Monthly Oil Production 2km" = "monthly_oil_2km",
  "Monthly Gas Production 2km" = "monthly_gas_2km",
  "Distance to WRP" = "dist_wrp",
  "WRP Capacity" = "closest_wrp_capacity",
  "Angle to WRP" = "angle_wrp",
  "Distance to Dominguez Channel" = "dist_dc",
  "Elevation" = "elevation",
  "Enhanced Vegetation Index" = "EVI",
  "Number of Daily Odor Complaints" = "num_odor_complaints",
  "Odor Event" = "disaster",
  "Year" = "year",
  "Hour of Day (sin)" = "hour_sin",
  "Hour of Day (cos)" = "hour_cos",
  "Month of Year (sin)" = "month_sin",
  "Month of Year (cos)" = "month_cos",
  "Day of Week (sin)" = "weekday_sin",
  "Day of Week (cos)" = "weekday_cos",
  "Day of Year (sin)" = "julian_sin",
  "Day of Year (cos)" = "julian_cos"
)
```

```{r}
hourly_predictors <- c('year',
                       'month_sin', 'month_cos',
                       'hour_sin', 'hour_cos',
                       'weekday_sin', 'weekday_cos',
                       'julian_sin', 'julian_cos',
                       'wd_sin', 'wd_cos', 'ws_avg', 
                       'hourly_downwind_ref', 'dist_wrp', 'dist_ref',
                       'mon_utm_x', 'mon_utm_y', 'monthly_oil_2km', 'monthly_gas_2km', 
                       'active_2km', 'inactive_2km', 'hourly_downwind_wrp', 
                       'elevation', 'EVI', 'num_odor_complaints',
                       'dist_dc', 'closest_wrp_capacity', 'hourly_temp', 'hourly_hum', 'hourly_precip') 

da_predictors <- c('year',
                   'month_sin', 'month_cos',
                   'hour_sin', 'hour_cos',
                   'weekday_sin', 'weekday_cos',
                   'julian_sin', 'julian_cos',
                   'daily_wd_sin', 'daily_wd_cos', 'daily_ws', 
                   'daily_downwind_ref', 'dist_wrp', 'dist_ref',
                   'mon_utm_x', 'mon_utm_y', 'monthly_oil_2km', 'monthly_gas_2km', 
                   'active_2km', 'inactive_2km', 'daily_downwind_wrp', 'elevation', 'EVI', 'num_odor_complaints',
                   'dist_dc', 'closest_wrp_capacity', 'daily_temp', 'daily_hum', 'daily_precip') 

dm_predictors <- c('year',
                   'month_sin', 'month_cos',
                   'hour_sin', 'hour_cos',
                   'weekday_sin', 'weekday_cos',
                   'julian_sin', 'julian_cos',
                   'hourly_wd_sin', 'hourly_wd_cos', 'hourly_ws', 
                   'hourly_downwind_ref', 'dist_wrp', 'dist_ref',
                   'mon_utm_x', 'mon_utm_y', 'monthly_oil_2km', 'monthly_gas_2km', 
                   'active_2km', 'inactive_2km', 'hourly_downwind_wrp', 'elevation', 'EVI', 'num_odor_complaints',
                   'dist_dc', 'closest_wrp_capacity', 'hourly_temp', 'hourly_hum', 'hourly_precip') 

rename_dist <- c('dist_ref' = 'inv_dist_ref_p2',
                 'dist_dc' = 'inv_dist_dc_p2',
                 'dist_wrp' = 'inv_dist_wrp_p2')

### Train matrices
hourly_excl_dis_train_Xmat <- hourly_excl_dis_train %>%
  rename(rename_dist) %>%
  select(all_of(hourly_predictors)) %>%
  as.matrix()

hourly_full_train_Xmat <-
  hourly_full_train %>%
  rename(rename_dist) %>%
  select(all_of(hourly_predictors)) %>%
  as.matrix()

daily_daha_excl_dis_train_Xmat <-
  daily_ha_excl_dis_train %>%
  rename(rename_dist) %>%
  select(all_of(da_predictors)) %>%
  as.matrix()

daily_daha_full_train_Xmat <-
  daily_ha_full_train %>%
  rename(rename_dist) %>%
  select(all_of(da_predictors)) %>%
  as.matrix()

daily_dmha_excl_dis_train_Xmat <-
  daily_ha_excl_dis_train %>%
  rename(rename_dist) %>%
  select(all_of(dm_predictors)) %>%
  as.matrix()

daily_dmha_full_train_Xmat <-
  daily_ha_full_train %>%
  rename(rename_dist) %>%
  select(all_of(dm_predictors)) %>%
  as.matrix()

### Test matrices
hourly_excl_dis_test_Xmat <- hourly_excl_dis_test %>%
  rename(rename_dist) %>%
  select(all_of(hourly_predictors)) %>%
  as.matrix()

hourly_full_test_Xmat <-
  hourly_full_test %>%
  rename(rename_dist) %>%
  select(all_of(hourly_predictors)) %>%
  as.matrix()

daily_daha_excl_dis_test_Xmat <-
  daily_ha_excl_dis_test %>%
  rename(rename_dist) %>%
  select(all_of(da_predictors)) %>%
  as.matrix()

daily_daha_full_test_Xmat <-
  daily_ha_full_test %>%
  rename(rename_dist) %>%
  select(all_of(da_predictors)) %>%
  as.matrix()


daily_dmha_excl_dis_test_Xmat <-
  daily_ha_excl_dis_test %>%
  rename(rename_dist) %>%
  select(all_of(dm_predictors)) %>%
  as.matrix()

daily_dmha_full_test_Xmat <-
  daily_ha_full_test %>%
  rename(rename_dist) %>%
  select(all_of(dm_predictors)) %>%
  as.matrix()

### Q99 matrices
daily_dmha_full_test_for_q99_Xmat <- 
  daily_ha_full_test %>%
  rename(rename_dist) %>%
  select(all_of(c(dm_predictors, 'day')))

daily_dmha_full_full_for_q99_Xmat <- 
  daily_ha_full_train %>%
  bind_rows(daily_ha_full_test) %>%
  rename(rename_dist) %>%
  select(all_of(c(dm_predictors, 'day')))

ha_full_full_for_q99_Xmat <- 
  hourly_full_train %>%
  bind_rows(hourly_full_test) %>%
  rename(rename_dist) %>%
  mutate(origin = as.Date("2020-01-01")) %>%
  select(all_of(c(hourly_predictors, 'day', 'dayhour')))

monitor_coord_map <- hourly_train %>% select(Monitor, mon_utm_x, mon_utm_y) %>% distinct()
```

```{r}
######## Save some files for QXGB #######
# write_parquet(monitor_coord_map, '../../data/monitor_coord_map.parquet')
# write_parquet(ha_full_full_for_q99_Xmat %>% as_tibble(), 
#               '../../data/ha_full_full_for_py_v2.parquet')
# write_parquet(bind_rows(hourly_full_train %>% select(H2S_hourly_avg),
#                         hourly_full_test %>% select(H2S_hourly_avg)),
#               '../../data/ha_full_full_y_for_py_v2.parquet')
############
```

```{r}
get_test_metrics <- function(model, test_mat, y_test, transformation = 'None',
                             lag_mean = NULL){
  y_pred <- predict(model, newdata = test_mat)
  if (!is.null(lag_mean)) y_pred <- y_pred + lag_mean
  R2 <- R2(y_pred, y_test)
  RMSE <- RMSE(y_pred, y_test)
  coef <- summary(lm(y_test ~ y_pred))$coefficients[2, 1]
  if (transformation == 'log'){
    BT_R2 <- R2(exp(y_pred), exp(y_test))
    BT_RMSE <- RMSE(exp(y_pred), exp(y_test))
    BT_coef <- summary(lm(exp(y_test) ~ exp(y_pred)))$coefficients[2, 1]
  } else if (transformation == 'None') {
    BT_R2 <- NA
    BT_RMSE <- NA
    BT_coef <- NA
  }
  return(list(pred = y_pred, R2 = R2, RMSE = RMSE, coef = coef, 
              BT_R2 = BT_R2, BT_RMSE = BT_RMSE, BT_coef = BT_coef))
}

get_event_metrics <- function(disaster, y_pred, y_test, transformation = 'None'){
  # Compute event period vs non-event period stats
  if (!(is.null(disaster) | sum(disaster) == 0)) {
    disaster <- as.logical(disaster)
    event_y_pred <- y_pred[disaster]
    event_y_test <- y_test[disaster]
    normal_y_pred <- y_pred[!disaster]
    normal_y_test <- y_test[!disaster]
    
    event_R2 <- R2(event_y_pred, event_y_test)
    event_RMSE <- RMSE(event_y_pred, event_y_test)
    event_coef <- summary(lm(event_y_test ~ event_y_pred))$coefficients[2, 1]
    normal_R2 <- R2(normal_y_pred, normal_y_test)
    normal_RMSE <- RMSE(normal_y_pred, normal_y_test)
    normal_coef <- summary(lm(normal_y_test ~ normal_y_pred))$coefficients[2, 1]
    
    if (transformation == 'log'){
      BT_event_R2 <- R2(exp(event_y_pred), exp(event_y_test))
      BT_event_RMSE <- RMSE(exp(event_y_pred), exp(event_y_test))
      BT_event_coef <- summary(lm(exp(event_y_test) ~ exp(event_y_pred)))$coefficients[2, 1]
      BT_normal_R2 <- R2(exp(normal_y_pred), exp(normal_y_test))
      BT_normal_RMSE <- RMSE(exp(normal_y_pred), exp(normal_y_test))
      BT_normal_coef <- summary(lm(exp(normal_y_test) ~ exp(normal_y_pred)))$coefficients[2, 1]
    } else if (transformation == 'None') {
      BT_event_R2 <- NA
      BT_event_RMSE <- NA
      BT_event_coef <- NA
      BT_normal_R2 <- NA
      BT_normal_RMSE <- NA
      BT_normal_coef <- NA
    }
    
  } else {
    for (met in c('R2', 'RMSE', 'coef')) {
      for (period in c('event', 'normal')) {
        for (t in c('', 'BT'))
          if (t == ''){
            assign(paste(period, met, sep='_'), NA)
          } else {
            assign(paste(t, period, met, sep='_'), NA)
          }
        
      }
    }
  }

  return(list(event_R2 = event_R2, event_RMSE = event_RMSE, event_coef = event_coef, 
              normal_R2 = normal_R2, normal_RMSE = normal_RMSE, normal_coef = normal_coef,
              BT_event_R2 = BT_event_R2, BT_event_RMSE = BT_event_RMSE, BT_event_coef = BT_event_coef,
              BT_normal_R2 = BT_normal_R2, BT_normal_RMSE = BT_normal_RMSE, BT_normal_coef = BT_normal_coef))
}

get_val_metrics <- function(model, transformation = 'None'){
  if (transformation == 'None') {
    fold_stat <- model$pred %>% group_by(Resample) %>% 
    summarise(R2 = R2(pred, obs), RMSE = RMSE(pred, obs))
    R2 <- mean(fold_stat$R2)
    RMSE <- mean(fold_stat$RMSE)
    coef <- summary(lm(model$pred$obs ~ model$pred$pred))$coefficients[2, 1]
    BT_R2 <- NA
    BT_RMSE <- NA
    BT_coef <- NA
  } else if (transformation == 'log') {
    fold_stat <- model$pred %>% group_by(Resample) %>% 
      summarise(R2 = R2(pred, obs), RMSE = RMSE(pred, obs),
                R2_BT = R2(exp(pred), exp(obs)), RMSE_BT = RMSE(exp(pred), exp(obs)))
    R2 <- mean(fold_stat$R2)
    BT_R2 <- mean(fold_stat$R2_BT)
    RMSE <- mean(fold_stat$RMSE)
    BT_RMSE <- mean(fold_stat$RMSE_BT)
    coef <- summary(lm(model$pred$obs ~ model$pred$pred))$coefficients[2, 1]
    BT_coef <- summary(lm(exp(model$pred$obs) ~ exp(model$pred$pred)))$coefficients[2, 1]
  }

  return(list(R2 = R2, RMSE = RMSE, coef = coef, 
              BT_R2 = BT_R2, BT_RMSE = BT_RMSE, BT_coef = BT_coef))
}

get_test_result_row <- function(name, response, daterange, test_metrics, event_metrics){
  test_metrics_row <- as_tibble_row(test_metrics[-which(names(test_metrics)=='pred')])
  event_metrics_row <- as_tibble_row(event_metrics)
  result <- tibble_row(Model = name, model_response_name = response, daterange = daterange,
                       test_metrics_row, event_metrics_row)
  return(result)
}

get_val_result_row <- function(name, response, daterange, val_metrics){
  val_metrics_row <- as_tibble_row(val_metrics)
  result <- tibble_row(Model = name, model_response_name = response, daterange = daterange,
                       val_metrics_row)
  return(result)
}

# helper to fit and summarize
lm_label <- function(df) {
  fit <- lm(obs ~ pred, data = df)
  b0 <- coef(fit)[1]; b1 <- coef(fit)[2]
  r2 <- summary(fit)$r.squared
  rmse <- sqrt(mean((df$pred - df$obs)^2, na.rm = TRUE))
  sprintf("y=%.2f+%.2fx, R²=%.2f, RMSE=%.2f", b0, b1, r2, rmse)
}

get_ovp_plot <- function(obs, pred, BT = FALSE,
                         disaster = NULL, anno_y = Inf, anno_x = -Inf) {
  if (!BT) {
    plot_df <- tibble(obs = obs,  pred = pred)
  } else {
    plot_df <- tibble(obs = exp(obs),  pred = exp(pred))
  }
  
  if (!is.null(disaster)) {
    plot_df$disaster <- as.numeric(disaster)
  } else {
    plot_df$disaster <- 0
  }
  
  plot <- ggplot(plot_df, aes(x = pred, y = obs, color = factor(disaster)))
  
  if (is.null(disaster)) {
    # Get label
    lab_txt <- lm_label(plot_df)
    label_df <- tibble(x = -Inf, y = Inf, lab = lab_txt)
  } else {
    lab_txt_all <- lm_label(plot_df)
    lab_txt_event <- lm_label(plot_df %>% filter(disaster == 1))
    lab_txt_routine <- lm_label(plot_df %>% filter(disaster == 0))
    lab_txt <- str_glue("All: {lab_txt_all}\nRoutine: {lab_txt_routine}\nEvent: {lab_txt_event}")
  }
  
  plot <- plot +
    geom_point(alpha = 0.9, show.legend = FALSE) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
    geom_smooth(aes(group = 1), method = "lm", se = FALSE, color = "black") +
    scale_color_manual(values = c(`0` = "dodgerblue", `1` = "tomato")) +
    labs(y = 'Observed (ppb)', x = 'Predicted (ppb)') +
    annotate(
      "label",
      x = -Inf, y = Inf,
      hjust = 0, vjust = 1,
      label = lab_txt,
      size = 4,
      fill = "white",
      color = "black",
      label.size = 0,      
      alpha = 0.85
    ) +
    theme_bw()
  
  return(plot)
}
```

```{r}
get_response_base_name <- function(response) {
  if (str_detect(response, 'daha')) {
    return('H2S_daily_avg_from_hourly_avg')
  } else if (str_detect(response, 'dmha')) {
    return('H2S_daily_max_from_hourly_avg')
  } else if (str_detect(response, 'ha')) {
    return('H2S_hourly_avg')
  }
}

get_observed <- function(data, 
                         response = c('daha', 'dmha', 'ha',
                                      'daha_log', 'dmha_log', 'ha_log')) {
  obs <- data %>% pull(get_response_base_name(response))
  
  if (str_detect(response, 'log')) {
    obs <- log(obs)
  }
  
  return(obs)
}

# For 80/20 tests with new variables but existing hyperparameter tunings.
compute_weights <- function(y_train, weight_q = 0.95, weight_max = 10){
  # Compute weights: Weight obs above 95th-percentile higher
  # weight_q = 0.95, weight_max = 20
  q    <- quantile(y_train, weight_q, na.rm = TRUE)
  ymax <- max(y_train, na.rm = TRUE)
  den  <- ymax - q
  w <- 1 + (weight_max - 1) * pmin(1, pmax(0, (y_train - q) / den))
  return(w)
}

fit_new_8020 <- function(model, X_train, y_train, 
                         weight_q = 0.95, weight_max = 10){
  # y_train should be already transformed
  
  # Compute weights
  w <- compute_weights(y_train, weight_q, weight_max)  
  
  # Get design matrix
  dtrain <- xgb.DMatrix(X_train, label = y_train, weight = w)
  
  set.seed(42)
  new_model <- xgb.train(
    params = as.list(model$bestTune[names(model$bestTune) != 'nrounds']),
    data = dtrain,
    nrounds = model$bestTune$nrounds,
    watchlist = list(train = dtrain),
    verbose = 0
  )
  
  return(new_model)
}


control_single <- trainControl(
  method = "none",
  verboseIter=TRUE
)

get_predictors <- function(response, lag = FALSE) {
  response_base <- str_replace(response, '_log', '')
  
  if (response_base == 'daha') {
    predictors <- da_predictors
  } else if (response_base == 'dmha') {
    predictors <- dm_predictors
  } else if (response_base == 'ha') {
    predictors <- hourly_predictors
  }
  
  if (lag == TRUE) predictors <- c(predictors, 'mean')
  
  return(predictors)
}

# Spatial Cross Validation
spatial_validation <- function(model, 
                               data, 
                               name,
                               response = c('daha', 'dmha', 'ha',
                                            'daha_log', 'dmha_log', 'ha_log'),
                               daterange,
                               lag = FALSE,
                               weight_q = 0.95,
                               weight_max = 10) {
  data <- data %>%
    rename(rename_dist)
  
  # Remove Chico
  data <- data %>% filter(Monitor != 'Chico')
  
  response_name <- get_response_base_name(response)
  transformation <- ifelse(str_detect(response, 'log'), 'log', 'None')
  
  predictors <- get_predictors(response, lag)
  
  # Data should be the whole data set for the model
  cv_table <- tibble(
    Model = character(),
    model_response_name = character(),
    daterange = character(),
    Monitor = character(),
    "R2" = numeric(),
    "RMSE" = numeric(),
    "coef" = numeric(),
    "event_R2" = numeric(),
    "event_RMSE" = numeric(),
    "event_coef" = numeric(),
    "normal_R2" = numeric(),
    "normal_RMSE" = numeric(),
    "normal_coef" = numeric()
  )
  cv_plots <- list()
  monitors <- unique(data$Monitor)
  for (monitor in monitors) {
    print(str_glue("Running LOSO model: leaving out {monitor}..."))
    # Prep the data sets (Leave-one-monitor out)
    train <- data %>% filter(Monitor != monitor)
    test <- data %>% filter(Monitor == monitor)
    
    # Y(s)
    y_train <- get_observed(train, response)
    y_test <- get_observed(test, response)
      
    if (lag == TRUE){
      # Add the mean response of the k-1 monitors as a predictor
      # Compute daily means Y(s') or hourly means
      if (response %in% c('ha', 'ha_log')) {
        means <- train %>%
          mutate(response = if (transformation == 'log') 
            log(.data[[response_name]]) else .data[[response_name]]) %>%
          group_by(dayhour) %>%
          summarise(mean = mean(response))
        train <- train %>%
          left_join(means, join_by(dayhour))
        test <- test %>%
          left_join(means, join_by(dayhour))
      } else {
        means <- train %>%
          mutate(response = if (transformation == 'log') 
            log(.data[[response_name]]) else .data[[response_name]]) %>%
          group_by(day) %>%
          summarise(mean = mean(response))
        train <- train %>%
          left_join(means, join_by(day))
        test <- test %>%
          left_join(means, join_by(day))
      }
      
      X_train <- train %>% select(all_of(predictors)) %>%
        mutate(mean = coalesce(mean, 0)) %>%
        as.matrix()
      X_test <- test %>% select(all_of(predictors)) %>%
        mutate(mean = coalesce(mean, 0)) %>%
        as.matrix()
    
      # Then, target is Y(s)-Y(s') instead of Y(s)
      # XGBoost implementation
      y_train <- y_train-X_train[,'mean']
      # y_test <- y_test-X_test[,'mean']
      # Below is used for caret
      # train <- X_train %>% mutate(response = y_train-mean)
      # test <- X_test %>% mutate(response = y_test-mean)
      
      # Compute weights on lagged scale (high deviation -> high weight)
      w <- compute_weights(y_train, weight_q, weight_max)  
    } else {
      X_train <- train %>% select(all_of(predictors)) %>% as.matrix()
      X_test <- test %>% select(all_of(predictors)) %>% as.matrix()
      # Below is used for caret
      # train <- X_train %>% mutate(response = y_train)
      # test <- X_test %>% mutate(response = y_test)
      
      # Compute weights on original scale
      w <- compute_weights(y_train, weight_q, weight_max)  
    }
    
    # Get design matrix
    dtrain <- xgb.DMatrix(X_train, label = y_train, weight = w)
    
    set.seed(42)
    # Fit model without this site with xgboost + weight
    lomo_model <- xgb.train(
      params = as.list(model$bestTune[names(model$bestTune) != 'nrounds']),
      data = dtrain,
      nrounds = model$bestTune$nrounds,
      watchlist = list(train = dtrain),
      verbose = 0
    )
    
    # Fit model without this site with caret
    # lomo_model <- train(response~.,
    #                            method = 'xgbTree',
    #                            data = train,
    #                            trControl = control_single,
    #                            tuneGrid = model$bestTune,
    #                            importance=TRUE, 
    #                            verbosity = 0, 
    #                            verbose=FALSE)
    
    # Get testing metrics on the left out site
    if (lag) lag_mean <- X_test[,'mean'] else lag_mean <- NULL
    test_metrics <- 
      get_test_metrics(lomo_model, X_test, y_test, transformation = transformation,
                       lag_mean = lag_mean)
    
    y_test_pred <- test_metrics$pred
  
    disaster_idx <- test$year == 2021 & test$month %in% 10:12
    event_metrics <- get_event_metrics(disaster_idx, y_test_pred, y_test, transformation)
    
    if (transformation == 'log') {
      test_metrics <- test_metrics[str_detect(names(test_metrics), 'BT')]
      names(test_metrics) <- str_replace(names(test_metrics), 'BT_', '')
      event_metrics <- event_metrics[str_detect(names(event_metrics), 'BT')]
      names(event_metrics) <- str_replace(names(event_metrics), 'BT_', '')
      
    } else {
      test_metrics <- test_metrics[!str_detect(names(test_metrics), 'BT')]
      event_metrics <- event_metrics[!str_detect(names(event_metrics), 'BT')]
    }
    
    if (sum(disaster_idx) == 0) disaster_idx <- NULL
    plot <- get_ovp_plot(y_test, y_test_pred, BT = transformation == 'log',
                         disaster = disaster_idx)
    
    plot <- plot +
      labs(title = paste0('Left out monitor: ', monitor))
    
    cv_plots[[monitor]] <- plot
    
    test_metrics_row <- as_tibble_row(test_metrics[c('R2', 'RMSE', 'coef')])
    event_metrics_row <- as_tibble_row(event_metrics)
    row <- tibble_row(Model = name, model_response_name = response, daterange = daterange,
                      test_metrics_row, event_metrics_row, Monitor = monitor)
    cv_table <- cv_table %>%
      add_row(row)
  }
  
  result <- list(cv_table = cv_table, cv_plots = cv_plots)
  return(result)
}

temporal_validation <- function(model, 
                               data, 
                               name,
                               response = c('daha', 'dmha', 'ha',
                                            'daha_log', 'dmha_log', 'ha_log'),
                               daterange,
                               weight_q = 0.95,
                               weight_max = 10) {
  data <- data %>%
    rename(rename_dist)
  
  response_name <- get_response_base_name(response)
  transformation <- ifelse(str_detect(response, 'log'), 'log', 'None')
  
  # Remove July/Aug 2021 for training
  train <- data %>% filter(!(year==2021 & month %in% 7:8)) 
  test <- data %>% filter(year==2021 & month %in% 7:8)
  
  y_train <- get_observed(train, response)
  y_test <- get_observed(test, response)
  
  # Compute weights on non-lagged scale
  w <- compute_weights(y_train, weight_q, weight_max)  
    
  # Select predictors
  predictors <- get_predictors(response)
  X_train <- train %>% select(all_of(predictors)) %>% as.matrix()
  X_test <- test %>% select(all_of(predictors)) %>% as.matrix()
  
  # Below is for xgboost implementation
  # Get design matrix
  dtrain <- xgb.DMatrix(X_train, label = y_train, weight = w)

  set.seed(42)
  # Fit model without this site with xgboost + weight
  loeo_model <- xgb.train(
    params = as.list(model$bestTune[names(model$bestTune) != 'nrounds']),
    data = dtrain,
    nrounds = model$bestTune$nrounds,
    watchlist = list(train = dtrain),
    verbose = 0
  )
    
  # Below is for caret implementation
  # train <- X_train %>% mutate(response = y_train)
  # test <- X_test %>% mutate(response = y_test)
  # 
  # # Fit model without the small event
  # loeo_model <- train(response~.,
  #                     method = 'xgbTree',
  #                     data = train,
  #                     trControl = control_single,
  #                     tuneGrid = model$bestTune,
  #                     importance=TRUE, 
  #                     verbosity = 0, 
  #                     verbose=FALSE)
  
  # Get testing metrics on the left out site
  test_metrics <- 
    get_test_metrics(loeo_model, X_test, y_test, transformation = transformation)
  
  y_test_pred <- test_metrics$pred
  
  if (transformation == 'log') {
    test_metrics <- test_metrics[str_detect(names(test_metrics), 'BT')]
    names(test_metrics) <- str_replace(names(test_metrics), 'BT_', '')
  } else {
    test_metrics <- test_metrics[!str_detect(names(test_metrics), 'BT')]
  }
  
  plot <- get_ovp_plot(y_test, y_test_pred, BT = transformation == 'log')
  
  test_metrics_row <- as_tibble_row(test_metrics[c('R2', 'RMSE', 'coef')])
  row <- tibble_row(Model = name, model_response_name = response, daterange = daterange,
                    test_metrics_row)
  
  return(list(result = row, plot = plot))
}

fit_final_model <- function(model, 
                            data, 
                            response = c('daha_log', 'dmha_log')) {
  data <- data %>%
    rename(rename_dist)
  
  response_name <- get_response_base_name(response)
  transformation <- ifelse(str_detect(response, 'log'), 'log', 'None')
  
  predictors <- get_predictors(response, lag = TRUE)
  
  # Data should be the whole data set for the model

  # Prep the data sets
  train <- data
  
  # Y(s)
  y_train <- get_observed(train, response)
    
  # ADD LAG
  # Add the mean response of the k-1 monitors as a predictor
  # Compute daily means Y(s') or hourly means
  if (response %in% c('ha', 'ha_log')) {
    means <- train %>%
      mutate(response = if (transformation == 'log') 
        log(.data[[response_name]]) else .data[[response_name]]) %>%
      group_by(dayhour) %>%
      summarise(mean = mean(response))
    train <- train %>%
      left_join(means, join_by(dayhour))
  } else {
    means <- train %>%
      mutate(response = if (transformation == 'log') 
        log(.data[[response_name]]) else .data[[response_name]]) %>%
      group_by(day) %>%
      summarise(mean = mean(response))
    train <- train %>%
      left_join(means, join_by(day))
  }
  
  X_train <- train %>% select(all_of(predictors)) %>%
    mutate(mean = coalesce(mean, 0)) %>%
    as.matrix()

  # Then, target is Y(s)-Y(s') instead of Y(s)
  # XGBoost implementation
  y_train <- y_train-X_train[,'mean']
  
  # Get design matrix
  dtrain <- xgb.DMatrix(X_train, label = y_train)
  
  set.seed(42)
  # Fit model without this site with xgboost + weight
  final_model <- xgb.train(
    params = as.list(model$bestTune[names(model$bestTune) != 'nrounds']),
    data = dtrain,
    nrounds = model$bestTune$nrounds,
    watchlist = list(train = dtrain),
    verbose = 0
  )

  return(final_model)
}
```

```{r}
hourly_responses <- c('H2S_hourly_avg', 'log(H2S_hourly_avg)')
daily_from_hour_avg_responses <- c('H2S_daily_avg_from_hourly_avg', 
                                   'log(H2S_daily_avg_from_hourly_avg)',
                                   'H2S_daily_max_from_hourly_avg', 
                                   'log(H2S_daily_max_from_hourly_avg)')

get_data <- function(response, daterange, type) {
  if (response %in% hourly_responses) {
      data <- get(paste0('hourly_', daterange, '_', type))
  } else if (response %in% daily_from_hour_avg_responses) {
      data <- get(paste0('daily_ha_', daterange, '_', type))
  }
  
  return(data)
}

get_curr_data <- function(response = c('daha', 'dmha', 'ha',
                                       'daha_log', 'dmha_log', 'ha_log'), 
                          daterange = c('excl_dis', 'full_unstrat')){
  
  daterange <- str_replace(daterange, '_unstrat', '')
  transformation <- ifelse(str_detect(response, 'log'), 'log', 'None')
  response_base <- str_replace(response, '_log', '')
  
  # Get X
  if (response_base == 'ha') {
    X_train <- get(paste('hourly', daterange, 'train', 'Xmat', sep = '_'))
    X_test <- get(paste('hourly', daterange, 'test', 'Xmat', sep = '_'))
  } else {
    X_train <- get(paste('daily', response_base, daterange, 'train', 'Xmat', sep = '_'))
    X_test <- get(paste('daily', response_base, daterange, 'test', 'Xmat', sep = '_'))
  }
  
  
  # Get y
  response_colname <- get_response_base_name(response)
  y_train <- get_data(response_colname, daterange, 'train') %>%
    pull(response_colname)
  y_test <- get_data(response_colname, daterange, 'test') %>%
    pull(response_colname)
  
  if (transformation == 'log') {
    y_train <- log(y_train)
    y_test <- log(y_test)
  }
  return(list(X_train = X_train, X_test = X_test, y_train = y_train, y_test = y_test))
}

get_fi_names <- function(response) {
  response_base <- str_replace(response, '_log', '')
  response_names <- get(paste0(response_base, '_', 'names'))
  full_names <- c(fixed_names, response_names)
  return(full_names)
}

get_fi_plot <- function(model, names, nfeatures = 15){
  # This is for caret
  # imp<-varImp(model,scale=FALSE)
  
  # This is for xgb.Booster
  imp <- xgb.importance(model = model)
  
  # rename variables
  imp$Feature <- ifelse(
    imp$Feature %in% unname(names),
    names(names)[match(imp$Feature, unname(names))],
    imp$Feature
  )

  imp <- tibble(variable = imp$Feature, importance = imp$Gain)
  
  plot <- imp %>%
    top_n(nfeatures, importance) %>%
    ggplot(aes(x=reorder(variable, importance), y=importance)) + 
    geom_point() +
    geom_segment(aes(x=variable,xend=variable,y=0,yend=importance)) +
    coord_flip() +
    theme_light() +
    theme(axis.text.y = element_text(size = 10),
          axis.title.x=element_blank(),
          axis.title.y=element_blank())
  return(plot)
}

get_test_dis_idx <- function(response, daterange) {
  daterange <- str_replace(daterange, '_unstrat', '')
  response_colname <- get_response_base_name(response)
  data <- get_data(response_colname, daterange, 'test')
  disaster_idx <- data$year == 2021 & data$month %in% 10:12
  if (sum(disaster_idx) == 0) disaster_idx <- NULL
  return(disaster_idx)
}
```



## DAHA {.tabset .tabset-pills}
```{r}
response <- 'daha'
transformation <- ifelse(str_detect(response, 'log'), 'log', 'None')
```

### Exclude Event {.tabset .tabset-pills} 
#### 80/20 test
```{r echo = T, results = 'hide'}
daterange <- 'excl_dis'
filename <- paste0('fit.xgb_', response, '_', daterange, '.rds')
assign(paste0('fit.xgb_', response, '_', daterange), readRDS(paste0(model_folder, filename)))
paste0('fit.xgb_', response, '_', daterange)
```

```{r}
v8020_weight_max <- 1
curr_model <- get(paste0('fit.xgb_', response, '_', daterange))
curr_data <- get_curr_data(response, daterange)
curr_X_train <- curr_data$X_train; curr_y_train <- curr_data$y_train
curr_X_test <- curr_data$X_test; curr_y_test <- curr_data$y_test
# Get new model (new variables + weight)
new_model <- fit_new_8020(curr_model, curr_X_train, curr_y_train, weight_max = v8020_weight_max)
```

```{r}
# Feature importance of new model
get_fi_plot(new_model, get_fi_names(response))
```

```{r}
test_metrics <- get_test_metrics(new_model, 
                                 curr_X_test,
                                 curr_y_test,
                                 transformation)
test_pred <- test_metrics$pred
disaster_idx <- get_test_dis_idx(response, daterange)
event_metrics <- get_event_metrics(disaster_idx, test_pred, curr_y_test, transformation)
```

```{r}
assign(paste(response, daterange, "ovp", sep = '_'),  
       get_ovp_plot(curr_y_test, test_pred, BT = transformation == 'log',
                                       disaster_idx))
get(paste(response, daterange, "ovp", sep = '_'))
```

```{r}
if (daterange == 'full_unstrat' | !(response  %in% c('daha', 'daha_log'))) {
  response_base <- str_replace(response, '_log', '')
  lim <- case_when(response_base == 'daha' ~ 4,
                    response_base == 'dmha' ~ 10,
                    response_base == 'ha' ~ 10,
                    .default = 4)
  assign(paste(response, daterange, "subset", lim, "ovp", sep = '_'),  
         get(paste(response, daterange, "ovp", sep = '_')) + 
           coord_cartesian(xlim = c(0, lim), ylim = c(0, lim)))
  print(get(paste(response, daterange, "subset", lim, "ovp", sep = '_')))
}
```


```{r}
test_result <- test_result %>%
  add_row(get_test_result_row('Exclude Event', response, daterange,
                              test_metrics, event_metrics))

negative_pred <- negative_pred %>%
  add_row(tibble_row(Model = 'Exclude Event', 
                     model_response_name = response, 
                     daterange = daterange,
                     'negative' = sum(test_pred < 0)))
```

#### Leave-one-out spatial tests {.tabset .tabset-pills}
##### With Lag

```{r}
data <- bind_rows(daily_ha_excl_dis_train, daily_ha_excl_dis_test)
lag_weight_max <- 1
# spatial_cv <- spatial_validation(curr_model, data, 'Exclude Event',
# response, daterange, lag = TRUE, weight_max = lag_weight_max)
#saveRDS(spatial_cv, paste0('../../tests/spatial_test/spatial_test_wlag', response, '_', daterange,  '_wgt', lag_weight_max, '.rds'))
spatial_cv <- readRDS(paste0('../../tests/spatial_test/spatial_test_wlag', response, '_', daterange,  '_wgt', lag_weight_max, '.rds'))
spatial_cv$cv_table
```

```{r}
for (i in seq_along(spatial_cv$cv_plots)) {
  print(spatial_cv$cv_plots[[i]])
}
```

```{r, eval = TRUE}
p<-wrap_plots(spatial_cv$cv_plots[1:14], ncol = 3)
p
```

```{r}
spatial_avg_result <- spatial_cv$cv_table %>%
  mutate(R2_mean = mean(c(event_R2, normal_R2), na.rm=TRUE),
         RMSE_mean = mean(c(event_RMSE, normal_RMSE), na.rm=TRUE),
         coef_mean = mean(c(event_coef, normal_coef), na.rm=TRUE)) %>%
  mutate(R2_mean = coalesce(R2_mean, R2),
         RMSE_mean =  coalesce(event_RMSE, RMSE),
         coef_mean =  coalesce(coef_mean, coef))

spatial_avg_result <- spatial_avg_result %>% 
  select(where(~is.numeric(.x))) %>% 
  colMeans(na.rm=TRUE)

spatial_test_result <- spatial_test_result %>%
  add_row(bind_cols(as_tibble_row(c(Model = 'Exclude Event',
                                    model_response_name = response,
                                    daterange = daterange, lag = TRUE, coordinates = TRUE)),
          as_tibble_row(spatial_avg_result)))
spatial_avg_result
```

```{r}
w <- data %>%
  group_by(Monitor) %>%
  summarise(n = n())
spatial_table_w <- spatial_cv$cv_table %>%
  left_join(w, join_by("Monitor"))

spatial_table_w <- spatial_table_w %>%
  mutate(R2_mean = mean(c(event_R2, normal_R2), na.rm=TRUE),
         RMSE_mean = mean(c(event_RMSE, normal_RMSE), na.rm=TRUE),
         coef_mean = mean(c(event_coef, normal_coef), na.rm=TRUE)) %>%
  mutate(R2_mean = coalesce(R2_mean, R2),
         RMSE_mean =  coalesce(event_RMSE, RMSE),
         coef_mean =  coalesce(coef_mean, coef))

spatial_weighted_avg_result <- spatial_table_w %>% 
  select(where(~is.numeric(.x))) %>% 
  summarise(across(!all_of("n") , ~ weighted.mean(.x, n, na.rm = TRUE)))

spatial_weighted_test_result <- spatial_weighted_test_result %>%
  add_row(bind_cols(as_tibble_row(c(Model = 'Exclude Event',
                                    model_response_name = response,
                                    daterange = daterange, lag = TRUE, coordinates = TRUE)),
          as_tibble_row(spatial_weighted_avg_result%>%unlist())))

spatial_weighted_avg_result
```


#### Leave-one-out temporal test
```{r}
temporal_weight_max <- 1
# temporal_test <- temporal_validation(curr_model, data, 'Exclude Event',
                                     # response, daterange, weight_max = temporal_weight_max)

# saveRDS(temporal_test, paste0('../../tests/temporal_test/temporal_test_', response, '_', daterange, '_wgt', temporal_weight_max,'.rds'))
temporal_test <- readRDS(paste0('../../tests/temporal_test/temporal_test_', response, '_', daterange, '_wgt', temporal_weight_max,'.rds'))

temporal_test_result <- temporal_test_result %>%
  add_row(temporal_test$result)

temporal_test$result
```

```{r}
temporal_test$plot
```


### Everything w.o Disaster Indicator {.tabset .tabset-pills}
#### 80/20 test
```{r echo = T, results = 'hide'}
daterange <- 'full_unstrat'
filename <- paste0('fit.xgb_', response, '_', daterange, '.rds')
assign(paste0('fit.xgb_', response, '_', daterange), readRDS(paste0(model_folder, filename)))
paste0('fit.xgb_', response, '_', daterange)
```
```{r}
curr_model <- get(paste0('fit.xgb_', response, '_', daterange))
curr_data <- get_curr_data(response, daterange)
curr_X_train <- curr_data$X_train; curr_y_train <- curr_data$y_train
curr_X_test <- curr_data$X_test; curr_y_test <- curr_data$y_test
# Get new model (new variables + weight)
new_model <- fit_new_8020(curr_model, curr_X_train, curr_y_train, weight_max = v8020_weight_max)
```

```{r}
# Feature importance of new model
get_fi_plot(new_model, get_fi_names(response))
```

```{r}
test_metrics <- get_test_metrics(new_model, 
                                 curr_X_test,
                                 curr_y_test,
                                 transformation)
test_pred <- test_metrics$pred
disaster_idx <- get_test_dis_idx(response, daterange)
event_metrics <- get_event_metrics(disaster_idx, test_pred, curr_y_test, transformation)
```

```{r}
assign(paste(response, daterange, "ovp", sep = '_'),  
       get_ovp_plot(curr_y_test, test_pred, BT = transformation == 'log',
                                       disaster_idx))
get(paste(response, daterange, "ovp", sep = '_'))
```

```{r}
if (daterange == 'full_unstrat' | !(response  %in% c('daha', 'daha_log'))) {
  response_base <- str_replace(response, '_log', '')
  lim <- case_when(response_base == 'daha' ~ 4,
                    response_base == 'dmha' ~ 10,
                    response_base == 'ha' ~ 10,
                    .default = 4)
  assign(paste(response, daterange, "subset", lim, "ovp", sep = '_'),  
         get(paste(response, daterange, "ovp", sep = '_')) + 
           coord_cartesian(xlim = c(0, lim), ylim = c(0, lim)))
  print(get(paste(response, daterange, "subset", lim, "ovp", sep = '_')))
}
```

```{r}
test_result_unstrat <- test_result_unstrat %>%
  add_row(get_test_result_row('Everything w.o Event Indicator', response, daterange,
                              test_metrics, event_metrics))

negative_pred <- negative_pred %>%
  add_row(tibble_row(Model = 'Everything w.o Event Indicator', 
                     model_response_name = response, 
                     daterange = daterange,
                     'negative' = sum(test_pred < 0)))
```

#### Leave-one-out spatial tests {.tabset .tabset-pills}
##### With Lag

```{r}
data <- bind_rows(daily_ha_full_train, daily_ha_full_test)
# spatial_cv <- spatial_validation(curr_model, data, 'Everything w.o Event Indicator',
# response, daterange, lag = TRUE, weight_max = lag_weight_max)
#saveRDS(spatial_cv, paste0('../../tests/spatial_test/spatial_test_wlag', response, '_', daterange,  '_wgt', lag_weight_max, '.rds'))
spatial_cv <- readRDS(paste0('../../tests/spatial_test/spatial_test_wlag', response, '_', daterange,  '_wgt', lag_weight_max, '.rds'))
spatial_cv$cv_table
```

```{r}
for (i in seq_along(spatial_cv$cv_plots)) {
  print(spatial_cv$cv_plots[[i]])
}
```

```{r, eval = TRUE}
p<-wrap_plots(spatial_cv$cv_plots[1:14], ncol = 3)
p
```

```{r}
spatial_avg_result <- spatial_cv$cv_table %>%
  mutate(R2_mean = mean(c(event_R2, normal_R2), na.rm=TRUE),
         RMSE_mean = mean(c(event_RMSE, normal_RMSE), na.rm=TRUE),
         coef_mean = mean(c(event_coef, normal_coef), na.rm=TRUE)) %>%
  mutate(R2_mean = coalesce(R2_mean, R2),
         RMSE_mean =  coalesce(event_RMSE, RMSE),
         coef_mean =  coalesce(coef_mean, coef))

spatial_avg_result <- spatial_avg_result %>% 
  select(where(~is.numeric(.x))) %>% 
  colMeans(na.rm=TRUE)

spatial_test_result <- spatial_test_result %>%
  add_row(bind_cols(as_tibble_row(c(Model = 'Everything w.o Event Indicator',
                                    model_response_name = response,
                                    daterange = daterange, lag = TRUE, coordinates = TRUE)),
          as_tibble_row(spatial_avg_result)))
spatial_avg_result
```

```{r}
w <- data %>%
  group_by(Monitor) %>%
  summarise(n = n())
spatial_table_w <- spatial_cv$cv_table %>%
  left_join(w, join_by("Monitor"))

spatial_table_w <- spatial_table_w %>%
  mutate(R2_mean = mean(c(event_R2, normal_R2), na.rm=TRUE),
         RMSE_mean = mean(c(event_RMSE, normal_RMSE), na.rm=TRUE),
         coef_mean = mean(c(event_coef, normal_coef), na.rm=TRUE)) %>%
  mutate(R2_mean = coalesce(R2_mean, R2),
         RMSE_mean =  coalesce(event_RMSE, RMSE),
         coef_mean =  coalesce(coef_mean, coef))

spatial_weighted_avg_result <- spatial_table_w %>% 
  select(where(~is.numeric(.x))) %>% 
  summarise(across(!all_of("n") , ~ weighted.mean(.x, n, na.rm = TRUE)))

spatial_weighted_test_result <- spatial_weighted_test_result %>%
  add_row(bind_cols(as_tibble_row(c(Model = 'Everything w.o Event Indicator',
                                    model_response_name = response,
                                    daterange = daterange, lag = TRUE, coordinates = TRUE)),
          as_tibble_row(spatial_weighted_avg_result%>%unlist())))

spatial_weighted_avg_result
```

#### Leave-one-out temporal test
```{r}
# temporal_test <- temporal_validation(curr_model, data, 'Everything w.o Event Indicator',
                                     # response, daterange, weight_max = temporal_weight_max)

# saveRDS(temporal_test, paste0('../../tests/temporal_test/temporal_test_', response, '_', daterange, '_wgt', temporal_weight_max,'.rds'))
temporal_test <- readRDS(paste0('../../tests/temporal_test/temporal_test_', response, '_', daterange, '_wgt', temporal_weight_max,'.rds'))

temporal_test_result <- temporal_test_result %>%
  add_row(temporal_test$result)

temporal_test$result
```

```{r}
temporal_test$plot
```

## Log DAHA {.tabset .tabset-pills}
```{r}
response <- 'daha_log'
transformation <- ifelse(str_detect(response, 'log'), 'log', 'None')
```

### Exclude Event {.tabset .tabset-pills}
#### 80/20 test
```{r}
daterange <- 'excl_dis'
filename <- paste0('fit.xgb_', response, '_', daterange, '.rds')
assign(paste0('fit.xgb_', response, '_', daterange), readRDS(paste0(model_folder, filename)))
paste0('fit.xgb_', response, '_', daterange)
```

```{r}
curr_model <- get(paste0('fit.xgb_', response, '_', daterange))
curr_data <- get_curr_data(response, daterange)
curr_X_train <- curr_data$X_train; curr_y_train <- curr_data$y_train
curr_X_test <- curr_data$X_test; curr_y_test <- curr_data$y_test
# Get new model (new variables + weight)
new_model <- fit_new_8020(curr_model, curr_X_train, curr_y_train, weight_max = v8020_weight_max)
```

```{r}
# Feature importance of new model
get_fi_plot(new_model, get_fi_names(response))
```

```{r}
test_metrics <- get_test_metrics(new_model, 
                                 curr_X_test,
                                 curr_y_test,
                                 transformation)
test_pred <- test_metrics$pred
disaster_idx <- get_test_dis_idx(response, daterange)
event_metrics <- get_event_metrics(disaster_idx, test_pred, curr_y_test, transformation)
```

```{r}
assign(paste(response, daterange, "ovp", sep = '_'),  
       get_ovp_plot(curr_y_test, test_pred, BT = transformation == 'log',
                                       disaster_idx))
get(paste(response, daterange, "ovp", sep = '_'))
```

```{r}
if (daterange == 'full_unstrat' | !(response  %in% c('daha', 'daha_log'))) {
  response_base <- str_replace(response, '_log', '')
  lim <- case_when(response_base == 'daha' ~ 4,
                    response_base == 'dmha' ~ 10,
                    response_base == 'ha' ~ 10,
                    .default = 4)
  assign(paste(response, daterange, "subset", lim, "ovp", sep = '_'),  
         get(paste(response, daterange, "ovp", sep = '_')) + 
           coord_cartesian(xlim = c(0, lim), ylim = c(0, lim)))
  print(get(paste(response, daterange, "subset", lim, "ovp", sep = '_')))
}
```


```{r}
test_result_unstrat <- test_result_unstrat %>%
  add_row(get_test_result_row('Exclude Event', response, daterange,
                              test_metrics, event_metrics))

negative_pred <- negative_pred %>%
  add_row(tibble_row(Model = 'Exclude Event', 
                     model_response_name = response, 
                     daterange = daterange,
                     'negative' = sum(test_pred < 0)))
```

#### Leave-one-out spatial tests {.tabset .tabset-pills}
##### With Lag

```{r}
data <- bind_rows(daily_ha_excl_dis_train, daily_ha_excl_dis_test)
# spatial_cv <- spatial_validation(curr_model, data, 'Exclude Event',
# response, daterange, lag = TRUE, weight_max = lag_weight_max)
#saveRDS(spatial_cv, paste0('../../tests/spatial_test/spatial_test_wlag', response, '_', daterange,  '_wgt', lag_weight_max, '.rds'))
spatial_cv <- readRDS(paste0('../../tests/spatial_test/spatial_test_wlag', response, '_', daterange,  '_wgt', lag_weight_max, '.rds'))
spatial_cv$cv_table
```

```{r}
for (i in seq_along(spatial_cv$cv_plots)) {
  print(spatial_cv$cv_plots[[i]])
}
```

```{r, eval = TRUE}
p<-wrap_plots(spatial_cv$cv_plots[1:14], ncol = 3)
p
```

```{r}
spatial_avg_result <- spatial_cv$cv_table %>%
  mutate(R2_mean = mean(c(event_R2, normal_R2), na.rm=TRUE),
         RMSE_mean = mean(c(event_RMSE, normal_RMSE), na.rm=TRUE),
         coef_mean = mean(c(event_coef, normal_coef), na.rm=TRUE)) %>%
  mutate(R2_mean = coalesce(R2_mean, R2),
         RMSE_mean =  coalesce(event_RMSE, RMSE),
         coef_mean =  coalesce(coef_mean, coef))

spatial_avg_result <- spatial_avg_result %>% 
  select(where(~is.numeric(.x))) %>% 
  colMeans(na.rm=TRUE)

spatial_test_result <- spatial_test_result %>%
  add_row(bind_cols(as_tibble_row(c(Model = 'Exclude Event',
                                    model_response_name = response,
                                    daterange = daterange, lag = TRUE, coordinates = TRUE)),
          as_tibble_row(spatial_avg_result)))
spatial_avg_result
```

```{r}
w <- data %>%
  group_by(Monitor) %>%
  summarise(n = n())
spatial_table_w <- spatial_cv$cv_table %>%
  left_join(w, join_by("Monitor"))

spatial_table_w <- spatial_table_w %>%
  mutate(R2_mean = mean(c(event_R2, normal_R2), na.rm=TRUE),
         RMSE_mean = mean(c(event_RMSE, normal_RMSE), na.rm=TRUE),
         coef_mean = mean(c(event_coef, normal_coef), na.rm=TRUE)) %>%
  mutate(R2_mean = coalesce(R2_mean, R2),
         RMSE_mean =  coalesce(event_RMSE, RMSE),
         coef_mean =  coalesce(coef_mean, coef))

spatial_weighted_avg_result <- spatial_table_w %>% 
  select(where(~is.numeric(.x))) %>% 
  summarise(across(!all_of("n") , ~ weighted.mean(.x, n, na.rm = TRUE)))

spatial_weighted_test_result <- spatial_weighted_test_result %>%
  add_row(bind_cols(as_tibble_row(c(Model = 'Exclude Event',
                                    model_response_name = response,
                                    daterange = daterange, lag = TRUE, coordinates = TRUE)),
          as_tibble_row(spatial_weighted_avg_result%>%unlist())))

spatial_weighted_avg_result
```

#### Leave-one-out temporal test
```{r}
# temporal_test <- temporal_validation(curr_model, data, 'Exclude Event',
                                     # response, daterange, weight_max = temporal_weight_max)

# saveRDS(temporal_test, paste0('../../tests/temporal_test/temporal_test_', response, '_', daterange, '_wgt', temporal_weight_max,'.rds'))
temporal_test <- readRDS(paste0('../../tests/temporal_test/temporal_test_', response, '_', daterange, '_wgt', temporal_weight_max,'.rds'))

temporal_test_result <- temporal_test_result %>%
  add_row(temporal_test$result)

temporal_test$result
```

```{r}
temporal_test$plot
```

### Everything w.o Disaster Indicator {.tabset .tabset-pills}
#### 80/20 test
```{r echo = T, results = 'hide'}
daterange <- 'full_unstrat'
filename <- paste0('fit.xgb_', response, '_', daterange, '.rds')
assign(paste0('fit.xgb_', response, '_', daterange), readRDS(paste0(model_folder, filename)))
paste0('fit.xgb_', response, '_', daterange)
```

```{r}
curr_model <- get(paste0('fit.xgb_', response, '_', daterange))
curr_data <- get_curr_data(response, daterange)
curr_X_train <- curr_data$X_train; curr_y_train <- curr_data$y_train
curr_X_test <- curr_data$X_test; curr_y_test <- curr_data$y_test
# Get new model (new variables + weight)
new_model <- fit_new_8020(curr_model, curr_X_train, curr_y_train, weight_max = v8020_weight_max)
```

```{r}
# Feature importance of new model
get_fi_plot(new_model, get_fi_names(response))
```

```{r}
test_metrics <- get_test_metrics(new_model, 
                                 curr_X_test,
                                 curr_y_test,
                                 transformation)
test_pred <- test_metrics$pred
disaster_idx <- get_test_dis_idx(response, daterange)
event_metrics <- get_event_metrics(disaster_idx, test_pred, curr_y_test, transformation)
```

```{r}
assign(paste(response, daterange, "ovp", sep = '_'),  
       get_ovp_plot(curr_y_test, test_pred, BT = transformation == 'log',
                                       disaster_idx))
get(paste(response, daterange, "ovp", sep = '_'))
```

```{r}
if (daterange == 'full_unstrat' | !(response  %in% c('daha', 'daha_log'))) {
  response_base <- str_replace(response, '_log', '')
  lim <- case_when(response_base == 'daha' ~ 4,
                    response_base == 'dmha' ~ 10,
                    response_base == 'ha' ~ 10,
                    .default = 4)
  assign(paste(response, daterange, "subset", lim, "ovp", sep = '_'),  
         get(paste(response, daterange, "ovp", sep = '_')) + 
           coord_cartesian(xlim = c(0, lim), ylim = c(0, lim)))
  print(get(paste(response, daterange, "subset", lim, "ovp", sep = '_')))
}
```


```{r}
test_result_unstrat <- test_result_unstrat %>%
  add_row(get_test_result_row('Everything w.o Event Indicator', response, daterange,
                              test_metrics, event_metrics))
```

#### Leave-one-out spatial tests {.tabset .tabset-pills}
##### With Lag

```{r}
data <- bind_rows(daily_ha_full_train, daily_ha_full_test)
# spatial_cv <- spatial_validation(curr_model, data, 'Everything w.o Event Indicator',
# response, daterange, lag = TRUE, weight_max = lag_weight_max)
#saveRDS(spatial_cv, paste0('../../tests/spatial_test/spatial_test_wlag', response, '_', daterange,  '_wgt', lag_weight_max, '.rds'))
spatial_cv <- readRDS(paste0('../../tests/spatial_test/spatial_test_wlag', response, '_', daterange,  '_wgt', lag_weight_max, '.rds'))
spatial_cv$cv_table
```

```{r}
for (i in seq_along(spatial_cv$cv_plots)) {
  print(spatial_cv$cv_plots[[i]])
}
```

```{r, eval = TRUE}
p<-wrap_plots(spatial_cv$cv_plots[1:14], ncol = 3)
p
```

```{r}
spatial_avg_result <- spatial_cv$cv_table %>%
  mutate(R2_mean = mean(c(event_R2, normal_R2), na.rm=TRUE),
         RMSE_mean = mean(c(event_RMSE, normal_RMSE), na.rm=TRUE),
         coef_mean = mean(c(event_coef, normal_coef), na.rm=TRUE)) %>%
  mutate(R2_mean = coalesce(R2_mean, R2),
         RMSE_mean =  coalesce(event_RMSE, RMSE),
         coef_mean =  coalesce(coef_mean, coef))

spatial_avg_result <- spatial_avg_result %>% 
  select(where(~is.numeric(.x))) %>% 
  colMeans(na.rm=TRUE)

spatial_test_result <- spatial_test_result %>%
  add_row(bind_cols(as_tibble_row(c(Model = 'Everything w.o Event Indicator',
                                    model_response_name = response,
                                    daterange = daterange, lag = TRUE, coordinates = TRUE)),
          as_tibble_row(spatial_avg_result)))
spatial_avg_result
```

```{r}
w <- data %>%
  group_by(Monitor) %>%
  summarise(n = n())
spatial_table_w <- spatial_cv$cv_table %>%
  left_join(w, join_by("Monitor"))

spatial_table_w <- spatial_table_w %>%
  mutate(R2_mean = mean(c(event_R2, normal_R2), na.rm=TRUE),
         RMSE_mean = mean(c(event_RMSE, normal_RMSE), na.rm=TRUE),
         coef_mean = mean(c(event_coef, normal_coef), na.rm=TRUE)) %>%
  mutate(R2_mean = coalesce(R2_mean, R2),
         RMSE_mean =  coalesce(event_RMSE, RMSE),
         coef_mean =  coalesce(coef_mean, coef))

spatial_weighted_avg_result <- spatial_table_w %>% 
  select(where(~is.numeric(.x))) %>% 
  summarise(across(!all_of("n") , ~ weighted.mean(.x, n, na.rm = TRUE)))

spatial_weighted_test_result <- spatial_weighted_test_result %>%
  add_row(bind_cols(as_tibble_row(c(Model = 'Everything w.o Event Indicator',
                                    model_response_name = response,
                                    daterange = daterange, lag = TRUE, coordinates = TRUE)),
          as_tibble_row(spatial_weighted_avg_result%>%unlist())))

spatial_weighted_avg_result
```

#### Leave-one-out temporal test
```{r}
# temporal_test <- temporal_validation(curr_model, data, 'Everything w.o Event Indicator',
                                     # response, daterange, weight_max = temporal_weight_max)

# saveRDS(temporal_test, paste0('../../tests/temporal_test/temporal_test_', response, '_', daterange, '_wgt', temporal_weight_max,'.rds'))
temporal_test <- readRDS(paste0('../../tests/temporal_test/temporal_test_', response, '_', daterange, '_wgt', temporal_weight_max,'.rds'))

temporal_test_result <- temporal_test_result %>%
  add_row(temporal_test$result)

temporal_test$result
```

```{r}
temporal_test$plot
```

```{r}
daha_log_full_loeo_plot <- temporal_test$plot
```

#### Fit final model
```{r}
log_daha_full_xgb <- fit_final_model(curr_model, data, response)
saveRDS(log_daha_full_xgb, paste0(model_folder, 'log_daha_full_final.rds'))
```

```{r}
# Save the daily lags for QXGB testing
# daily_lag <- data %>%
#   group_by(day) %>%
#   summarise(avg_da = mean(H2S_daily_avg_from_hourly_avg),
#             avg_dm = mean(H2S_daily_max_from_hourly_avg),
#             avg_log_da = mean(log(H2S_daily_avg_from_hourly_avg)),
#             avg_log_dm = mean(log(H2S_daily_max_from_hourly_avg)))
# saveRDS(daily_lag, '../../data/daily_lag.rds')
```


## DMHA {.tabset .tabset-pills}
```{r}
response <- 'dmha'
transformation <- ifelse(str_detect(response, 'log'), 'log', 'None')
```

### Exclude Event {.tabset .tabset-pills}
#### 80/20 test
```{r echo = T, results = 'hide'}
daterange <- 'excl_dis'
filename <- paste0('fit.xgb_', response, '_', daterange, '.rds')
assign(paste0('fit.xgb_', response, '_', daterange), readRDS(paste0(model_folder, filename)))
paste0('fit.xgb_', response, '_', daterange)
```

```{r}
curr_model <- get(paste0('fit.xgb_', response, '_', daterange))
curr_data <- get_curr_data(response, daterange)
curr_X_train <- curr_data$X_train; curr_y_train <- curr_data$y_train
curr_X_test <- curr_data$X_test; curr_y_test <- curr_data$y_test
# Get new model (new variables + weight)
new_model <- fit_new_8020(curr_model, curr_X_train, curr_y_train, weight_max = v8020_weight_max)
```

```{r}
# Feature importance of new model
get_fi_plot(new_model, get_fi_names(response))
```

```{r}
test_metrics <- get_test_metrics(new_model, 
                                 curr_X_test,
                                 curr_y_test,
                                 transformation)
test_pred <- test_metrics$pred
disaster_idx <- get_test_dis_idx(response, daterange)
event_metrics <- get_event_metrics(disaster_idx, test_pred, curr_y_test, transformation)
```

```{r}
assign(paste(response, daterange, "ovp", sep = '_'),  
       get_ovp_plot(curr_y_test, test_pred, BT = transformation == 'log',
                                       disaster_idx))
get(paste(response, daterange, "ovp", sep = '_'))
```

```{r}
if (daterange == 'full_unstrat' | !(response  %in% c('daha', 'daha_log'))) {
  response_base <- str_replace(response, '_log', '')
  lim <- case_when(response_base == 'daha' ~ 4,
                    response_base == 'dmha' ~ 10,
                    response_base == 'ha' ~ 10,
                    .default = 4)
  assign(paste(response, daterange, "subset", lim, "ovp", sep = '_'),  
         get(paste(response, daterange, "ovp", sep = '_')) + 
           coord_cartesian(xlim = c(0, lim), ylim = c(0, lim)))
  print(get(paste(response, daterange, "subset", lim, "ovp", sep = '_')))
}
```

```{r}
test_result <- test_result %>%
  add_row(get_test_result_row('Exclude Event', response, daterange,
                              test_metrics, event_metrics))

negative_pred <- negative_pred %>%
  add_row(tibble_row(Model = 'Exclude Event', 
                     model_response_name = response, 
                     daterange = daterange,
                     'negative' = sum(test_pred < 0)))
```

#### Leave-one-out spatial tests {.tabset .tabset-pills}
##### With Lag

```{r}
data <- bind_rows(daily_ha_excl_dis_train, daily_ha_excl_dis_test)
# spatial_cv <- spatial_validation(curr_model, data, 'Exclude Event',
# response, daterange, lag = TRUE, weight_max = lag_weight_max)
#saveRDS(spatial_cv, paste0('../../tests/spatial_test/spatial_test_wlag', response, '_', daterange,  '_wgt', lag_weight_max, '.rds'))
spatial_cv <- readRDS(paste0('../../tests/spatial_test/spatial_test_wlag', response, '_', daterange,  '_wgt', lag_weight_max, '.rds'))
spatial_cv$cv_table
```

```{r}
for (i in seq_along(spatial_cv$cv_plots)) {
  print(spatial_cv$cv_plots[[i]])
}
```

```{r, eval = TRUE}
p<-wrap_plots(spatial_cv$cv_plots[1:14], ncol = 3)
p
```

```{r}
spatial_avg_result <- spatial_cv$cv_table %>%
  mutate(R2_mean = mean(c(event_R2, normal_R2), na.rm=TRUE),
         RMSE_mean = mean(c(event_RMSE, normal_RMSE), na.rm=TRUE),
         coef_mean = mean(c(event_coef, normal_coef), na.rm=TRUE)) %>%
  mutate(R2_mean = coalesce(R2_mean, R2),
         RMSE_mean =  coalesce(event_RMSE, RMSE),
         coef_mean =  coalesce(coef_mean, coef))

spatial_avg_result <- spatial_avg_result %>% 
  select(where(~is.numeric(.x))) %>% 
  colMeans(na.rm=TRUE)

spatial_test_result <- spatial_test_result %>%
  add_row(bind_cols(as_tibble_row(c(Model = 'Exclude Event',
                                    model_response_name = response,
                                    daterange = daterange, lag = TRUE, coordinates = TRUE)),
          as_tibble_row(spatial_avg_result)))
spatial_avg_result
```

```{r}
w <- data %>%
  group_by(Monitor) %>%
  summarise(n = n())
spatial_table_w <- spatial_cv$cv_table %>%
  left_join(w, join_by("Monitor"))

spatial_table_w <- spatial_table_w %>%
  mutate(R2_mean = mean(c(event_R2, normal_R2), na.rm=TRUE),
         RMSE_mean = mean(c(event_RMSE, normal_RMSE), na.rm=TRUE),
         coef_mean = mean(c(event_coef, normal_coef), na.rm=TRUE)) %>%
  mutate(R2_mean = coalesce(R2_mean, R2),
         RMSE_mean =  coalesce(event_RMSE, RMSE),
         coef_mean =  coalesce(coef_mean, coef))

spatial_weighted_avg_result <- spatial_table_w %>% 
  select(where(~is.numeric(.x))) %>% 
  summarise(across(!all_of("n") , ~ weighted.mean(.x, n, na.rm = TRUE)))

spatial_weighted_test_result <- spatial_weighted_test_result %>%
  add_row(bind_cols(as_tibble_row(c(Model = 'Exclude Event',
                                    model_response_name = response,
                                    daterange = daterange, lag = TRUE, coordinates = TRUE)),
          as_tibble_row(spatial_weighted_avg_result%>%unlist())))

spatial_weighted_avg_result
```

#### Leave-one-out temporal test
```{r}
# temporal_test <- temporal_validation(curr_model, data, 'Exclude Event',
                                     # response, daterange, weight_max = temporal_weight_max)

# saveRDS(temporal_test, paste0('../../tests/temporal_test/temporal_test_', response, '_', daterange, '_wgt', temporal_weight_max,'.rds'))
temporal_test <- readRDS(paste0('../../tests/temporal_test/temporal_test_', response, '_', daterange, '_wgt', temporal_weight_max,'.rds'))

temporal_test_result <- temporal_test_result %>%
  add_row(temporal_test$result)

temporal_test$result
```

```{r}
temporal_test$plot
```


### Everything w.o Disaster Indicator {.tabset .tabset-pills}
#### 80/20 test
```{r echo = T, results = 'hide'}
daterange <- 'full_unstrat'
filename <- paste0('fit.xgb_', response, '_', daterange, '.rds')
assign(paste0('fit.xgb_', response, '_', daterange), readRDS(paste0(model_folder, filename)))
paste0('fit.xgb_', response, '_', daterange)
```

```{r}
curr_model <- get(paste0('fit.xgb_', response, '_', daterange))
curr_data <- get_curr_data(response, daterange)
curr_X_train <- curr_data$X_train; curr_y_train <- curr_data$y_train
curr_X_test <- curr_data$X_test; curr_y_test <- curr_data$y_test
# Get new model (new variables + weight)
new_model <- fit_new_8020(curr_model, curr_X_train, curr_y_train, weight_max = v8020_weight_max)
```

```{r}
# Feature importance of new model
get_fi_plot(new_model, get_fi_names(response))
```

```{r}
test_metrics <- get_test_metrics(new_model, 
                                 curr_X_test,
                                 curr_y_test,
                                 transformation)
test_pred <- test_metrics$pred
disaster_idx <- get_test_dis_idx(response, daterange)
event_metrics <- get_event_metrics(disaster_idx, test_pred, curr_y_test, transformation)
```

```{r}
assign(paste(response, daterange, "ovp", sep = '_'),  
       get_ovp_plot(curr_y_test, test_pred, BT = transformation == 'log',
                                       disaster_idx))
get(paste(response, daterange, "ovp", sep = '_'))
```

```{r}
if (daterange == 'full_unstrat' | !(response  %in% c('daha', 'daha_log'))) {
  response_base <- str_replace(response, '_log', '')
  lim <- case_when(response_base == 'daha' ~ 4,
                    response_base == 'dmha' ~ 10,
                    response_base == 'ha' ~ 10,
                    .default = 4)
  assign(paste(response, daterange, "subset", lim, "ovp", sep = '_'),  
         get(paste(response, daterange, "ovp", sep = '_')) + 
           coord_cartesian(xlim = c(0, lim), ylim = c(0, lim)))
  print(get(paste(response, daterange, "subset", lim, "ovp", sep = '_')))
}
```

```{r}
test_result_unstrat <- test_result_unstrat %>%
  add_row(get_test_result_row('Everything w.o Event Indicator', response, daterange,
                              test_metrics, event_metrics))


negative_pred <- negative_pred %>%
  add_row(tibble_row(Model = 'Everything w.o Event Indicator', 
                     model_response_name = response, 
                     daterange = daterange,
                     'negative' = sum(test_pred < 0)))
```

#### Leave-one-out spatial tests {.tabset .tabset-pills}
##### With Lag

```{r}
data <- bind_rows(daily_ha_full_train, daily_ha_full_test)
# spatial_cv <- spatial_validation(curr_model, data, 'Everything w.o Event Indicator',
# response, daterange, lag = TRUE, weight_max = lag_weight_max)
#saveRDS(spatial_cv, paste0('../../tests/spatial_test/spatial_test_wlag', response, '_', daterange,  '_wgt', lag_weight_max, '.rds'))
spatial_cv <- readRDS(paste0('../../tests/spatial_test/spatial_test_wlag', response, '_', daterange,  '_wgt', lag_weight_max, '.rds'))
spatial_cv$cv_table
```

```{r}
for (i in seq_along(spatial_cv$cv_plots)) {
  print(spatial_cv$cv_plots[[i]])
}
```

```{r, eval = TRUE}
p<-wrap_plots(spatial_cv$cv_plots[1:14], ncol = 3)
p
```

```{r}
spatial_avg_result <- spatial_cv$cv_table %>%
  mutate(R2_mean = mean(c(event_R2, normal_R2), na.rm=TRUE),
         RMSE_mean = mean(c(event_RMSE, normal_RMSE), na.rm=TRUE),
         coef_mean = mean(c(event_coef, normal_coef), na.rm=TRUE)) %>%
  mutate(R2_mean = coalesce(R2_mean, R2),
         RMSE_mean =  coalesce(event_RMSE, RMSE),
         coef_mean =  coalesce(coef_mean, coef))

spatial_avg_result <- spatial_avg_result %>% 
  select(where(~is.numeric(.x))) %>% 
  colMeans(na.rm=TRUE)

spatial_test_result <- spatial_test_result %>%
  add_row(bind_cols(as_tibble_row(c(Model = 'Everything w.o Event Indicator',
                                    model_response_name = response,
                                    daterange = daterange, lag = TRUE, coordinates = TRUE)),
          as_tibble_row(spatial_avg_result)))
spatial_avg_result
```

```{r}
w <- data %>%
  group_by(Monitor) %>%
  summarise(n = n())
spatial_table_w <- spatial_cv$cv_table %>%
  left_join(w, join_by("Monitor"))

spatial_table_w <- spatial_table_w %>%
  mutate(R2_mean = mean(c(event_R2, normal_R2), na.rm=TRUE),
         RMSE_mean = mean(c(event_RMSE, normal_RMSE), na.rm=TRUE),
         coef_mean = mean(c(event_coef, normal_coef), na.rm=TRUE)) %>%
  mutate(R2_mean = coalesce(R2_mean, R2),
         RMSE_mean =  coalesce(event_RMSE, RMSE),
         coef_mean =  coalesce(coef_mean, coef))

spatial_weighted_avg_result <- spatial_table_w %>% 
  select(where(~is.numeric(.x))) %>% 
  summarise(across(!all_of("n") , ~ weighted.mean(.x, n, na.rm = TRUE)))

spatial_weighted_test_result <- spatial_weighted_test_result %>%
  add_row(bind_cols(as_tibble_row(c(Model = 'Everything w.o Event Indicator',
                                    model_response_name = response,
                                    daterange = daterange, lag = TRUE, coordinates = TRUE)),
          as_tibble_row(spatial_weighted_avg_result%>%unlist())))

spatial_weighted_avg_result
```

#### Leave-one-out temporal test
```{r}
# temporal_test <- temporal_validation(curr_model, data, 'Everything w.o Event Indicator',
                                     # response, daterange, weight_max = temporal_weight_max)

# saveRDS(temporal_test, paste0('../../tests/temporal_test/temporal_test_', response, '_', daterange, '_wgt', temporal_weight_max,'.rds'))
temporal_test <- readRDS(paste0('../../tests/temporal_test/temporal_test_', response, '_', daterange, '_wgt', temporal_weight_max,'.rds'))

temporal_test_result <- temporal_test_result %>%
  add_row(temporal_test$result)

temporal_test$result
```

```{r}
temporal_test$plot
```

## Log DMHA {.tabset .tabset-pills}
```{r}
response <- 'dmha_log'
transformation <- ifelse(str_detect(response, 'log'), 'log', 'None')
```

### Exclude Event {.tabset .tabset-pills}
#### 80/20 test
```{r}
daterange <- 'excl_dis'
filename <- paste0('fit.xgb_', response, '_', daterange, '.rds')
assign(paste0('fit.xgb_', response, '_', daterange), readRDS(paste0(model_folder, filename)))
paste0('fit.xgb_', response, '_', daterange)
```

```{r}
curr_model <- get(paste0('fit.xgb_', response, '_', daterange))
curr_data <- get_curr_data(response, daterange)
curr_X_train <- curr_data$X_train; curr_y_train <- curr_data$y_train
curr_X_test <- curr_data$X_test; curr_y_test <- curr_data$y_test
# Get new model (new variables + weight)
new_model <- fit_new_8020(curr_model, curr_X_train, curr_y_train, weight_max = v8020_weight_max)
```

```{r}
# Feature importance of new model
get_fi_plot(new_model, get_fi_names(response))
```

```{r}
test_metrics <- get_test_metrics(new_model, 
                                 curr_X_test,
                                 curr_y_test,
                                 transformation)
test_pred <- test_metrics$pred
disaster_idx <- get_test_dis_idx(response, daterange)
event_metrics <- get_event_metrics(disaster_idx, test_pred, curr_y_test, transformation)
```

```{r}
assign(paste(response, daterange, "ovp", sep = '_'),  
       get_ovp_plot(curr_y_test, test_pred, BT = transformation == 'log',
                                       disaster_idx))
get(paste(response, daterange, "ovp", sep = '_'))
```

```{r}
if (daterange == 'full_unstrat' | !(response  %in% c('daha', 'daha_log'))) {
  response_base <- str_replace(response, '_log', '')
  lim <- case_when(response_base == 'daha' ~ 4,
                    response_base == 'dmha' ~ 10,
                    response_base == 'ha' ~ 10,
                    .default = 4)
  assign(paste(response, daterange, "subset", lim, "ovp", sep = '_'),  
         get(paste(response, daterange, "ovp", sep = '_')) + 
           coord_cartesian(xlim = c(0, lim), ylim = c(0, lim)))
  print(get(paste(response, daterange, "subset", lim, "ovp", sep = '_')))
}
```


```{r}
test_result <- test_result %>%
  add_row(get_test_result_row('Exclude Event', response, daterange,
                              test_metrics, event_metrics))
```

#### Leave-one-out spatial tests {.tabset .tabset-pills}
##### With Lag

```{r}
data <- bind_rows(daily_ha_excl_dis_train, daily_ha_excl_dis_test)
# spatial_cv <- spatial_validation(curr_model, data, 'Exclude Event',
# response, daterange, lag = TRUE, weight_max = lag_weight_max)
#saveRDS(spatial_cv, paste0('../../tests/spatial_test/spatial_test_wlag', response, '_', daterange,  '_wgt', lag_weight_max, '.rds'))
spatial_cv <- readRDS(paste0('../../tests/spatial_test/spatial_test_wlag', response, '_', daterange,  '_wgt', lag_weight_max, '.rds'))
spatial_cv$cv_table
```

```{r}
for (i in seq_along(spatial_cv$cv_plots)) {
  print(spatial_cv$cv_plots[[i]])
}
```

```{r, eval = TRUE}
p<-wrap_plots(spatial_cv$cv_plots[1:14], ncol = 3)
p
```

```{r}
spatial_avg_result <- spatial_cv$cv_table %>%
  mutate(R2_mean = mean(c(event_R2, normal_R2), na.rm=TRUE),
         RMSE_mean = mean(c(event_RMSE, normal_RMSE), na.rm=TRUE),
         coef_mean = mean(c(event_coef, normal_coef), na.rm=TRUE)) %>%
  mutate(R2_mean = coalesce(R2_mean, R2),
         RMSE_mean =  coalesce(event_RMSE, RMSE),
         coef_mean =  coalesce(coef_mean, coef))

spatial_avg_result <- spatial_avg_result %>% 
  select(where(~is.numeric(.x))) %>% 
  colMeans(na.rm=TRUE)

spatial_test_result <- spatial_test_result %>%
  add_row(bind_cols(as_tibble_row(c(Model = 'Exclude Event',
                                    model_response_name = response,
                                    daterange = daterange, lag = TRUE, coordinates = TRUE)),
          as_tibble_row(spatial_avg_result)))
spatial_avg_result
```

```{r}
w <- data %>%
  group_by(Monitor) %>%
  summarise(n = n())
spatial_table_w <- spatial_cv$cv_table %>%
  left_join(w, join_by("Monitor"))

spatial_table_w <- spatial_table_w %>%
  mutate(R2_mean = mean(c(event_R2, normal_R2), na.rm=TRUE),
         RMSE_mean = mean(c(event_RMSE, normal_RMSE), na.rm=TRUE),
         coef_mean = mean(c(event_coef, normal_coef), na.rm=TRUE)) %>%
  mutate(R2_mean = coalesce(R2_mean, R2),
         RMSE_mean =  coalesce(event_RMSE, RMSE),
         coef_mean =  coalesce(coef_mean, coef))

spatial_weighted_avg_result <- spatial_table_w %>% 
  select(where(~is.numeric(.x))) %>% 
  summarise(across(!all_of("n") , ~ weighted.mean(.x, n, na.rm = TRUE)))

spatial_weighted_test_result <- spatial_weighted_test_result %>%
  add_row(bind_cols(as_tibble_row(c(Model = 'Exclude Event',
                                    model_response_name = response,
                                    daterange = daterange, lag = TRUE, coordinates = TRUE)),
          as_tibble_row(spatial_weighted_avg_result%>%unlist())))

spatial_weighted_avg_result
```

#### Leave-one-out temporal test
```{r}
# temporal_test <- temporal_validation(curr_model, data, 'Exclude Event',
                                     # response, daterange, weight_max = temporal_weight_max)

# saveRDS(temporal_test, paste0('../../tests/temporal_test/temporal_test_', response, '_', daterange, '_wgt', temporal_weight_max,'.rds'))
temporal_test <- readRDS(paste0('../../tests/temporal_test/temporal_test_', response, '_', daterange, '_wgt', temporal_weight_max,'.rds'))

temporal_test_result <- temporal_test_result %>%
  add_row(temporal_test$result)

temporal_test$result
```

```{r}
temporal_test$plot
```

### Everything w.o Disaster Indicator {.tabset .tabset-pills}
#### 80/20 test
```{r echo = T, results = 'hide'}
daterange <- 'full_unstrat'
filename <- paste0('fit.xgb_', response, '_', daterange, '.rds')
assign(paste0('fit.xgb_', response, '_', daterange), readRDS(paste0(model_folder, filename)))
paste0('fit.xgb_', response, '_', daterange)
```

```{r}
curr_model <- get(paste0('fit.xgb_', response, '_', daterange))
curr_data <- get_curr_data(response, daterange)
curr_X_train <- curr_data$X_train; curr_y_train <- curr_data$y_train
curr_X_test <- curr_data$X_test; curr_y_test <- curr_data$y_test
# Get new model (new variables + weight)
new_model <- fit_new_8020(curr_model, curr_X_train, curr_y_train, weight_max = v8020_weight_max)
```

```{r}
# Feature importance of new model
get_fi_plot(new_model, get_fi_names(response))
```

```{r}
test_metrics <- get_test_metrics(new_model, 
                                 curr_X_test,
                                 curr_y_test,
                                 transformation)
test_pred <- test_metrics$pred
disaster_idx <- get_test_dis_idx(response, daterange)
event_metrics <- get_event_metrics(disaster_idx, test_pred, curr_y_test, transformation)
```

```{r}
assign(paste(response, daterange, "ovp", sep = '_'),  
       get_ovp_plot(curr_y_test, test_pred, BT = transformation == 'log',
                                       disaster_idx))
get(paste(response, daterange, "ovp", sep = '_'))
```

```{r}
if (daterange == 'full_unstrat' | !(response  %in% c('daha', 'daha_log'))) {
  response_base <- str_replace(response, '_log', '')
  lim <- case_when(response_base == 'daha' ~ 4,
                    response_base == 'dmha' ~ 10,
                    response_base == 'ha' ~ 10,
                    .default = 4)
  assign(paste(response, daterange, "subset", lim, "ovp", sep = '_'),  
         get(paste(response, daterange, "ovp", sep = '_')) + 
           coord_cartesian(xlim = c(0, lim), ylim = c(0, lim)))
  print(get(paste(response, daterange, "subset", lim, "ovp", sep = '_')))
}
```

```{r}
test_result_unstrat <- test_result_unstrat %>%
  add_row(get_test_result_row('Everything w.o Event Indicator', response, daterange,
                              test_metrics, event_metrics))
```

#### Leave-one-out spatial tests {.tabset .tabset-pills}
##### With Lag

```{r}
data <- bind_rows(daily_ha_full_train, daily_ha_full_test)
# spatial_cv <- spatial_validation(curr_model, data, 'Everything w.o Event Indicator',
# response, daterange, lag = TRUE, weight_max = lag_weight_max)
#saveRDS(spatial_cv, paste0('../../tests/spatial_test/spatial_test_wlag', response, '_', daterange,  '_wgt', lag_weight_max, '.rds'))
spatial_cv <- readRDS(paste0('../../tests/spatial_test/spatial_test_wlag', response, '_', daterange,  '_wgt', lag_weight_max, '.rds'))
spatial_cv$cv_table
```

```{r}
for (i in seq_along(spatial_cv$cv_plots)) {
  print(spatial_cv$cv_plots[[i]])
}
```

```{r, eval = TRUE}
p<-wrap_plots(spatial_cv$cv_plots[1:14], ncol = 3)
p
```

```{r}
spatial_avg_result <- spatial_cv$cv_table %>%
  mutate(R2_mean = mean(c(event_R2, normal_R2), na.rm=TRUE),
         RMSE_mean = mean(c(event_RMSE, normal_RMSE), na.rm=TRUE),
         coef_mean = mean(c(event_coef, normal_coef), na.rm=TRUE)) %>%
  mutate(R2_mean = coalesce(R2_mean, R2),
         RMSE_mean =  coalesce(event_RMSE, RMSE),
         coef_mean =  coalesce(coef_mean, coef))

spatial_avg_result <- spatial_avg_result %>% 
  select(where(~is.numeric(.x))) %>% 
  colMeans(na.rm=TRUE)

spatial_test_result <- spatial_test_result %>%
  add_row(bind_cols(as_tibble_row(c(Model = 'Everything w.o Event Indicator',
                                    model_response_name = response,
                                    daterange = daterange, lag = TRUE, coordinates = TRUE)),
          as_tibble_row(spatial_avg_result)))
spatial_avg_result
```

```{r}
w <- data %>%
  group_by(Monitor) %>%
  summarise(n = n())
spatial_table_w <- spatial_cv$cv_table %>%
  left_join(w, join_by("Monitor"))

spatial_table_w <- spatial_table_w %>%
  mutate(R2_mean = mean(c(event_R2, normal_R2), na.rm=TRUE),
         RMSE_mean = mean(c(event_RMSE, normal_RMSE), na.rm=TRUE),
         coef_mean = mean(c(event_coef, normal_coef), na.rm=TRUE)) %>%
  mutate(R2_mean = coalesce(R2_mean, R2),
         RMSE_mean =  coalesce(event_RMSE, RMSE),
         coef_mean =  coalesce(coef_mean, coef))

spatial_weighted_avg_result <- spatial_table_w %>% 
  select(where(~is.numeric(.x))) %>% 
  summarise(across(!all_of("n") , ~ weighted.mean(.x, n, na.rm = TRUE)))

spatial_weighted_test_result <- spatial_weighted_test_result %>%
  add_row(bind_cols(as_tibble_row(c(Model = 'Everything w.o Event Indicator',
                                    model_response_name = response,
                                    daterange = daterange, lag = TRUE, coordinates = TRUE)),
          as_tibble_row(spatial_weighted_avg_result%>%unlist())))

spatial_weighted_avg_result
```

#### Leave-one-out temporal test
```{r}
# temporal_test <- temporal_validation(curr_model, data, 'Everything w.o Event Indicator',
                                     # response, daterange, weight_max = temporal_weight_max)

# saveRDS(temporal_test, paste0('../../tests/temporal_test/temporal_test_', response, '_', daterange, '_wgt', temporal_weight_max,'.rds'))
temporal_test <- readRDS(paste0('../../tests/temporal_test/temporal_test_', response, '_', daterange, '_wgt', temporal_weight_max,'.rds'))

temporal_test_result <- temporal_test_result %>%
  add_row(temporal_test$result)

temporal_test$result
```

```{r}
temporal_test$plot
```

```{r}
dmha_log_full_loeo_plot <- temporal_test$plot
```


#### Fit final model
```{r}
log_dmha_full_xgb <- fit_final_model(curr_model, data, response)
saveRDS(log_dmha_full_xgb, paste0(model_folder, 'log_dmha_full_final.rds'))
```

## HA {.tabset .tabset-pills}
```{r}
response <- 'ha'
transformation <- ifelse(str_detect(response, 'log'), 'log', 'None')
```

### Exclude Event {.tabset .tabset-pills}
#### 80/20 test
```{r echo = T, results = 'hide'}
daterange <- 'excl_dis'
filename <- paste0('fit.xgb_', response, '_', daterange, '.rds')
assign(paste0('fit.xgb_', response, '_', daterange), readRDS(paste0(model_folder, filename)))
paste0('fit.xgb_', response, '_', daterange)
```

```{r}
curr_model <- get(paste0('fit.xgb_', response, '_', daterange))
curr_data <- get_curr_data(response, daterange)
curr_X_train <- curr_data$X_train; curr_y_train <- curr_data$y_train
curr_X_test <- curr_data$X_test; curr_y_test <- curr_data$y_test
# Get new model (new variables + weight)
new_model <- fit_new_8020(curr_model, curr_X_train, curr_y_train, weight_max = v8020_weight_max)
```

```{r}
# Feature importance of new model
get_fi_plot(new_model, get_fi_names(response))
```

```{r}
test_metrics <- get_test_metrics(new_model, 
                                 curr_X_test,
                                 curr_y_test,
                                 transformation)
test_pred <- test_metrics$pred
disaster_idx <- get_test_dis_idx(response, daterange)
event_metrics <- get_event_metrics(disaster_idx, test_pred, curr_y_test, transformation)
```

```{r}
assign(paste(response, daterange, "ovp", sep = '_'),  
       get_ovp_plot(curr_y_test, test_pred, BT = transformation == 'log',
                                       disaster_idx))
get(paste(response, daterange, "ovp", sep = '_'))
```

```{r}
if (daterange == 'full_unstrat' | !(response  %in% c('daha', 'daha_log'))) {
  response_base <- str_replace(response, '_log', '')
  lim <- case_when(response_base == 'daha' ~ 4,
                    response_base == 'dmha' ~ 10,
                    response_base == 'ha' ~ 10,
                    .default = 4)
  assign(paste(response, daterange, "subset", lim, "ovp", sep = '_'),  
         get(paste(response, daterange, "ovp", sep = '_')) + 
           coord_cartesian(xlim = c(0, lim), ylim = c(0, lim)))
  print(get(paste(response, daterange, "subset", lim, "ovp", sep = '_')))
}
```

```{r}
test_result <- test_result %>%
  add_row(get_test_result_row('Exclude Event', response, daterange,
                              test_metrics, event_metrics))

negative_pred <- negative_pred %>%
  add_row(tibble_row(Model = 'Exclude Event', 
                     model_response_name = response, 
                     daterange = daterange,
                     'negative' = sum(test_pred < 0)))
```

#### Leave-one-out spatial tests {.tabset .tabset-pills}
##### With Lag

```{r}
data <- bind_rows(hourly_excl_dis_train, hourly_excl_dis_test)
# spatial_cv <- spatial_validation(curr_model, data, 'Exclude Event',
# response, daterange, lag = TRUE, weight_max = lag_weight_max)
#saveRDS(spatial_cv, paste0('../../tests/spatial_test/spatial_test_wlag', response, '_', daterange,  '_wgt', lag_weight_max, '.rds'))
spatial_cv <- readRDS(paste0('../../tests/spatial_test/spatial_test_wlag', response, '_', daterange,  '_wgt', lag_weight_max, '.rds'))
spatial_cv$cv_table
```

```{r}
for (i in seq_along(spatial_cv$cv_plots)) {
  print(spatial_cv$cv_plots[[i]])
}
```

```{r, eval = TRUE}
p<-wrap_plots(spatial_cv$cv_plots[1:14], ncol = 3)
p
```

```{r}
spatial_avg_result <- spatial_cv$cv_table %>%
  mutate(R2_mean = mean(c(event_R2, normal_R2), na.rm=TRUE),
         RMSE_mean = mean(c(event_RMSE, normal_RMSE), na.rm=TRUE),
         coef_mean = mean(c(event_coef, normal_coef), na.rm=TRUE)) %>%
  mutate(R2_mean = coalesce(R2_mean, R2),
         RMSE_mean =  coalesce(event_RMSE, RMSE),
         coef_mean =  coalesce(coef_mean, coef))

spatial_avg_result <- spatial_avg_result %>% 
  select(where(~is.numeric(.x))) %>% 
  colMeans(na.rm=TRUE)

spatial_test_result <- spatial_test_result %>%
  add_row(bind_cols(as_tibble_row(c(Model = 'Exclude Event',
                                    model_response_name = response,
                                    daterange = daterange, lag = TRUE, coordinates = TRUE)),
          as_tibble_row(spatial_avg_result)))
spatial_avg_result
```

```{r}
w <- data %>%
  group_by(Monitor) %>%
  summarise(n = n())
spatial_table_w <- spatial_cv$cv_table %>%
  left_join(w, join_by("Monitor"))

spatial_table_w <- spatial_table_w %>%
  mutate(R2_mean = mean(c(event_R2, normal_R2), na.rm=TRUE),
         RMSE_mean = mean(c(event_RMSE, normal_RMSE), na.rm=TRUE),
         coef_mean = mean(c(event_coef, normal_coef), na.rm=TRUE)) %>%
  mutate(R2_mean = coalesce(R2_mean, R2),
         RMSE_mean =  coalesce(event_RMSE, RMSE),
         coef_mean =  coalesce(coef_mean, coef))

spatial_weighted_avg_result <- spatial_table_w %>% 
  select(where(~is.numeric(.x))) %>% 
  summarise(across(!all_of("n") , ~ weighted.mean(.x, n, na.rm = TRUE)))

spatial_weighted_test_result <- spatial_weighted_test_result %>%
  add_row(bind_cols(as_tibble_row(c(Model = 'Exclude Event',
                                    model_response_name = response,
                                    daterange = daterange, lag = TRUE, coordinates = TRUE)),
          as_tibble_row(spatial_weighted_avg_result%>%unlist())))

spatial_weighted_avg_result
```

#### Leave-one-out temporal test
```{r}
# temporal_test <- temporal_validation(curr_model, data, 'Exclude Event',
                                     # response, daterange, weight_max = temporal_weight_max)

# saveRDS(temporal_test, paste0('../../tests/temporal_test/temporal_test_', response, '_', daterange, '_wgt', temporal_weight_max,'.rds'))
temporal_test <- readRDS(paste0('../../tests/temporal_test/temporal_test_', response, '_', daterange, '_wgt', temporal_weight_max,'.rds'))

temporal_test_result <- temporal_test_result %>%
  add_row(temporal_test$result)

temporal_test$result
```

```{r}
temporal_test$plot
```

### Everything w.o Disaster Indicator {.tabset .tabset-pills}
#### 80/20 test
```{r echo = T, results = 'hide'}
daterange <- 'full_unstrat'
filename <- paste0('fit.xgb_', response, '_', daterange, '.rds')
assign(paste0('fit.xgb_', response, '_', daterange), readRDS(paste0(model_folder, filename)))
paste0('fit.xgb_', response, '_', daterange)
```

```{r}
curr_model <- get(paste0('fit.xgb_', response, '_', daterange))
curr_data <- get_curr_data(response, daterange)
curr_X_train <- curr_data$X_train; curr_y_train <- curr_data$y_train
curr_X_test <- curr_data$X_test; curr_y_test <- curr_data$y_test
# Get new model (new variables + weight)
new_model <- fit_new_8020(curr_model, curr_X_train, curr_y_train, weight_max = v8020_weight_max)
```

```{r}
# Feature importance of new model
get_fi_plot(new_model, get_fi_names(response))
```

```{r}
test_metrics <- get_test_metrics(new_model, 
                                 curr_X_test,
                                 curr_y_test,
                                 transformation)
test_pred <- test_metrics$pred
disaster_idx <- get_test_dis_idx(response, daterange)
event_metrics <- get_event_metrics(disaster_idx, test_pred, curr_y_test, transformation)
```

```{r}
assign(paste(response, daterange, "ovp", sep = '_'),  
       get_ovp_plot(curr_y_test, test_pred, BT = transformation == 'log',
                                       disaster_idx))
get(paste(response, daterange, "ovp", sep = '_'))
```

```{r}
if (daterange == 'full_unstrat' | !(response  %in% c('daha', 'daha_log'))) {
  response_base <- str_replace(response, '_log', '')
  lim <- case_when(response_base == 'daha' ~ 4,
                    response_base == 'dmha' ~ 10,
                    response_base == 'ha' ~ 10,
                    .default = 4)
  assign(paste(response, daterange, "subset", lim, "ovp", sep = '_'),  
         get(paste(response, daterange, "ovp", sep = '_')) + 
           coord_cartesian(xlim = c(0, lim), ylim = c(0, lim)))
  print(get(paste(response, daterange, "subset", lim, "ovp", sep = '_')))
}
```

```{r}
test_result_unstrat <- test_result_unstrat %>%
  add_row(get_test_result_row('Everything w.o Event Indicator', response, daterange,
                              test_metrics, event_metrics))

negative_pred <- negative_pred %>%
  add_row(tibble_row(Model = 'Everything w.o Event Indicator', 
                     model_response_name = response, 
                     daterange = daterange,
                     'negative' = sum(test_pred < 0)))
```

#### Leave-one-out spatial tests {.tabset .tabset-pills}
##### With Lag

```{r}
data <- bind_rows(hourly_full_train, hourly_full_test)
# spatial_cv <- spatial_validation(curr_model, data, 'Everything w.o Event Indicator',
# response, daterange, lag = TRUE, weight_max = lag_weight_max)
#saveRDS(spatial_cv, paste0('../../tests/spatial_test/spatial_test_wlag', response, '_', daterange,  '_wgt', lag_weight_max, '.rds'))
spatial_cv <- readRDS(paste0('../../tests/spatial_test/spatial_test_wlag', response, '_', daterange,  '_wgt', lag_weight_max, '.rds'))
spatial_cv$cv_table
```

```{r}
for (i in seq_along(spatial_cv$cv_plots)) {
  print(spatial_cv$cv_plots[[i]])
}
```

```{r, eval = TRUE}
p<-wrap_plots(spatial_cv$cv_plots[1:14], ncol = 3)
p
```

```{r}
spatial_avg_result <- spatial_cv$cv_table %>%
  mutate(R2_mean = mean(c(event_R2, normal_R2), na.rm=TRUE),
         RMSE_mean = mean(c(event_RMSE, normal_RMSE), na.rm=TRUE),
         coef_mean = mean(c(event_coef, normal_coef), na.rm=TRUE)) %>%
  mutate(R2_mean = coalesce(R2_mean, R2),
         RMSE_mean =  coalesce(event_RMSE, RMSE),
         coef_mean =  coalesce(coef_mean, coef))

spatial_avg_result <- spatial_avg_result %>% 
  select(where(~is.numeric(.x))) %>% 
  colMeans(na.rm=TRUE)

spatial_test_result <- spatial_test_result %>%
  add_row(bind_cols(as_tibble_row(c(Model = 'Everything w.o Event Indicator',
                                    model_response_name = response,
                                    daterange = daterange, lag = TRUE, coordinates = TRUE)),
          as_tibble_row(spatial_avg_result)))
spatial_avg_result
```

```{r}
w <- data %>%
  group_by(Monitor) %>%
  summarise(n = n())
spatial_table_w <- spatial_cv$cv_table %>%
  left_join(w, join_by("Monitor"))

spatial_table_w <- spatial_table_w %>%
  mutate(R2_mean = mean(c(event_R2, normal_R2), na.rm=TRUE),
         RMSE_mean = mean(c(event_RMSE, normal_RMSE), na.rm=TRUE),
         coef_mean = mean(c(event_coef, normal_coef), na.rm=TRUE)) %>%
  mutate(R2_mean = coalesce(R2_mean, R2),
         RMSE_mean =  coalesce(event_RMSE, RMSE),
         coef_mean =  coalesce(coef_mean, coef))

spatial_weighted_avg_result <- spatial_table_w %>% 
  select(where(~is.numeric(.x))) %>% 
  summarise(across(!all_of("n") , ~ weighted.mean(.x, n, na.rm = TRUE)))

spatial_weighted_test_result <- spatial_weighted_test_result %>%
  add_row(bind_cols(as_tibble_row(c(Model = 'Everything w.o Event Indicator',
                                    model_response_name = response,
                                    daterange = daterange, lag = TRUE, coordinates = TRUE)),
          as_tibble_row(spatial_weighted_avg_result%>%unlist())))

spatial_weighted_avg_result
```

#### Leave-one-out temporal test
```{r}
# temporal_test <- temporal_validation(curr_model, data, 'Everything w.o Event Indicator',
                                     # response, daterange, weight_max = temporal_weight_max)

# saveRDS(temporal_test, paste0('../../tests/temporal_test/temporal_test_', response, '_', daterange, '_wgt', temporal_weight_max,'.rds'))
temporal_test <- readRDS(paste0('../../tests/temporal_test/temporal_test_', response, '_', daterange, '_wgt', temporal_weight_max,'.rds'))

temporal_test_result <- temporal_test_result %>%
  add_row(temporal_test$result)

temporal_test$result
```

```{r}
temporal_test$plot
```

## Log HA {.tabset .tabset-pills}
```{r}
response <- 'ha_log'
transformation <- ifelse(str_detect(response, 'log'), 'log', 'None')
```

### Exclude Event {.tabset .tabset-pills}
#### 80/20 test
```{r}
daterange <- 'excl_dis'
filename <- paste0('fit.xgb_', response, '_', daterange, '.rds')
assign(paste0('fit.xgb_', response, '_', daterange), readRDS(paste0(model_folder, filename)))
paste0('fit.xgb_', response, '_', daterange)
```

```{r}
curr_model <- get(paste0('fit.xgb_', response, '_', daterange))
curr_data <- get_curr_data(response, daterange)
curr_X_train <- curr_data$X_train; curr_y_train <- curr_data$y_train
curr_X_test <- curr_data$X_test; curr_y_test <- curr_data$y_test
# Get new model (new variables + weight)
new_model <- fit_new_8020(curr_model, curr_X_train, curr_y_train, weight_max = v8020_weight_max)
```

```{r}
# Feature importance of new model
get_fi_plot(new_model, get_fi_names(response))
```

```{r}
test_metrics <- get_test_metrics(new_model, 
                                 curr_X_test,
                                 curr_y_test,
                                 transformation)
test_pred <- test_metrics$pred
disaster_idx <- get_test_dis_idx(response, daterange)
event_metrics <- get_event_metrics(disaster_idx, test_pred, curr_y_test, transformation)
```

```{r}
assign(paste(response, daterange, "ovp", sep = '_'),  
       get_ovp_plot(curr_y_test, test_pred, BT = transformation == 'log',
                                       disaster_idx))
get(paste(response, daterange, "ovp", sep = '_'))
```

```{r}
if (daterange == 'full_unstrat' | !(response  %in% c('daha', 'daha_log'))) {
  response_base <- str_replace(response, '_log', '')
  lim <- case_when(response_base == 'daha' ~ 4,
                    response_base == 'dmha' ~ 10,
                    response_base == 'ha' ~ 10,
                    .default = 4)
  assign(paste(response, daterange, "subset", lim, "ovp", sep = '_'),  
         get(paste(response, daterange, "ovp", sep = '_')) + 
           coord_cartesian(xlim = c(0, lim), ylim = c(0, lim)))
  print(get(paste(response, daterange, "subset", lim, "ovp", sep = '_')))
}
```

```{r}
test_result <- test_result %>%
  add_row(get_test_result_row('Exclude Event', response, daterange,
                              test_metrics, event_metrics))
```

#### Leave-one-out spatial tests {.tabset .tabset-pills}
##### With Lag

```{r}
data <- bind_rows(hourly_excl_dis_train, hourly_excl_dis_test)
# spatial_cv <- spatial_validation(curr_model, data, 'Exclude Event',
# response, daterange, lag = TRUE, weight_max = lag_weight_max)
#saveRDS(spatial_cv, paste0('../../tests/spatial_test/spatial_test_wlag', response, '_', daterange,  '_wgt', lag_weight_max, '.rds'))
spatial_cv <- readRDS(paste0('../../tests/spatial_test/spatial_test_wlag', response, '_', daterange,  '_wgt', lag_weight_max, '.rds'))
spatial_cv$cv_table
```

```{r}
for (i in seq_along(spatial_cv$cv_plots)) {
  print(spatial_cv$cv_plots[[i]])
}
```

```{r, eval = TRUE}
p<-wrap_plots(spatial_cv$cv_plots[1:14], ncol = 3)
p
```

```{r}
spatial_avg_result <- spatial_cv$cv_table %>%
  mutate(R2_mean = mean(c(event_R2, normal_R2), na.rm=TRUE),
         RMSE_mean = mean(c(event_RMSE, normal_RMSE), na.rm=TRUE),
         coef_mean = mean(c(event_coef, normal_coef), na.rm=TRUE)) %>%
  mutate(R2_mean = coalesce(R2_mean, R2),
         RMSE_mean =  coalesce(event_RMSE, RMSE),
         coef_mean =  coalesce(coef_mean, coef))

spatial_avg_result <- spatial_avg_result %>% 
  select(where(~is.numeric(.x))) %>% 
  colMeans(na.rm=TRUE)

spatial_test_result <- spatial_test_result %>%
  add_row(bind_cols(as_tibble_row(c(Model = 'Exclude Event',
                                    model_response_name = response,
                                    daterange = daterange, lag = TRUE, coordinates = TRUE)),
          as_tibble_row(spatial_avg_result)))
spatial_avg_result
```

```{r}
w <- data %>%
  group_by(Monitor) %>%
  summarise(n = n())
spatial_table_w <- spatial_cv$cv_table %>%
  left_join(w, join_by("Monitor"))

spatial_table_w <- spatial_table_w %>%
  mutate(R2_mean = mean(c(event_R2, normal_R2), na.rm=TRUE),
         RMSE_mean = mean(c(event_RMSE, normal_RMSE), na.rm=TRUE),
         coef_mean = mean(c(event_coef, normal_coef), na.rm=TRUE)) %>%
  mutate(R2_mean = coalesce(R2_mean, R2),
         RMSE_mean =  coalesce(event_RMSE, RMSE),
         coef_mean =  coalesce(coef_mean, coef))

spatial_weighted_avg_result <- spatial_table_w %>% 
  select(where(~is.numeric(.x))) %>% 
  summarise(across(!all_of("n") , ~ weighted.mean(.x, n, na.rm = TRUE)))

spatial_weighted_test_result <- spatial_weighted_test_result %>%
  add_row(bind_cols(as_tibble_row(c(Model = 'Exclude Event',
                                    model_response_name = response,
                                    daterange = daterange, lag = TRUE, coordinates = TRUE)),
          as_tibble_row(spatial_weighted_avg_result%>%unlist())))

spatial_weighted_avg_result
```

#### Leave-one-out temporal test
```{r}
# temporal_test <- temporal_validation(curr_model, data, 'Exclude Event',
                                     # response, daterange, weight_max = temporal_weight_max)

# saveRDS(temporal_test, paste0('../../tests/temporal_test/temporal_test_', response, '_', daterange, '_wgt', temporal_weight_max,'.rds'))
temporal_test <- readRDS(paste0('../../tests/temporal_test/temporal_test_', response, '_', daterange, '_wgt', temporal_weight_max,'.rds'))

temporal_test_result <- temporal_test_result %>%
  add_row(temporal_test$result)

temporal_test$result
```

```{r}
temporal_test$plot
```

### Everything w.o Disaster Indicator {.tabset .tabset-pills}
#### 80/20 test
```{r echo = T, results = 'hide'}
daterange <- 'full_unstrat'
filename <- paste0('fit.xgb_', response, '_', daterange, '.rds')
assign(paste0('fit.xgb_', response, '_', daterange), readRDS(paste0(model_folder, filename)))
paste0('fit.xgb_', response, '_', daterange)
```

```{r}
curr_model <- get(paste0('fit.xgb_', response, '_', daterange))
curr_data <- get_curr_data(response, daterange)
curr_X_train <- curr_data$X_train; curr_y_train <- curr_data$y_train
curr_X_test <- curr_data$X_test; curr_y_test <- curr_data$y_test
# Get new model (new variables + weight)
new_model <- fit_new_8020(curr_model, curr_X_train, curr_y_train, weight_max = v8020_weight_max)
```

```{r}
# Feature importance of new model
get_fi_plot(new_model, get_fi_names(response))
```

```{r}
test_metrics <- get_test_metrics(new_model, 
                                 curr_X_test,
                                 curr_y_test,
                                 transformation)
test_pred <- test_metrics$pred
disaster_idx <- get_test_dis_idx(response, daterange)
event_metrics <- get_event_metrics(disaster_idx, test_pred, curr_y_test, transformation)
```

```{r}
assign(paste(response, daterange, "ovp", sep = '_'),  
       get_ovp_plot(curr_y_test, test_pred, BT = transformation == 'log',
                                       disaster_idx))
get(paste(response, daterange, "ovp", sep = '_'))
```

```{r}
if (daterange == 'full_unstrat' | !(response  %in% c('daha', 'daha_log'))) {
  response_base <- str_replace(response, '_log', '')
  lim <- case_when(response_base == 'daha' ~ 4,
                    response_base == 'dmha' ~ 10,
                    response_base == 'ha' ~ 10,
                    .default = 4)
  assign(paste(response, daterange, "subset", lim, "ovp", sep = '_'),  
         get(paste(response, daterange, "ovp", sep = '_')) + 
           coord_cartesian(xlim = c(0, lim), ylim = c(0, lim)))
  print(get(paste(response, daterange, "subset", lim, "ovp", sep = '_')))
}
```

```{r}
test_result_unstrat <- test_result_unstrat %>%
  add_row(get_test_result_row('Everything w.o Event Indicator', response, daterange,
                              test_metrics, event_metrics))
```
  
#### Leave-one-out spatial tests {.tabset .tabset-pills}
##### With Lag
```{r}
data <- bind_rows(hourly_full_train, hourly_full_test)
# spatial_cv <- spatial_validation(curr_model, data, 'Everything w.o Event Indicator',
# response, daterange, lag = TRUE, weight_max = lag_weight_max)
#saveRDS(spatial_cv, paste0('../../tests/spatial_test/spatial_test_wlag', response, '_', daterange,  '_wgt', lag_weight_max, '.rds'))
spatial_cv <- readRDS(paste0('../../tests/spatial_test/spatial_test_wlag', response, '_', daterange,  '_wgt', lag_weight_max, '.rds'))
spatial_cv$cv_table
```

```{r}
for (i in seq_along(spatial_cv$cv_plots)) {
  print(spatial_cv$cv_plots[[i]])
}
```

```{r, eval = TRUE}
p<-wrap_plots(spatial_cv$cv_plots[1:14], ncol = 3)
p
```

```{r}
spatial_avg_result <- spatial_cv$cv_table %>%
  mutate(R2_mean = mean(c(event_R2, normal_R2), na.rm=TRUE),
         RMSE_mean = mean(c(event_RMSE, normal_RMSE), na.rm=TRUE),
         coef_mean = mean(c(event_coef, normal_coef), na.rm=TRUE)) %>%
  mutate(R2_mean = coalesce(R2_mean, R2),
         RMSE_mean =  coalesce(event_RMSE, RMSE),
         coef_mean =  coalesce(coef_mean, coef))

spatial_avg_result <- spatial_avg_result %>% 
  select(where(~is.numeric(.x))) %>% 
  colMeans(na.rm=TRUE)

spatial_test_result <- spatial_test_result %>%
  add_row(bind_cols(as_tibble_row(c(Model = 'Everything w.o Event Indicator',
                                    model_response_name = response,
                                    daterange = daterange, lag = TRUE, coordinates = TRUE)),
          as_tibble_row(spatial_avg_result)))
spatial_avg_result
```

```{r}
w <- data %>%
  group_by(Monitor) %>%
  summarise(n = n())

spatial_table_w <- spatial_cv$cv_table %>%
  left_join(w, join_by("Monitor"))

spatial_table_w <- spatial_table_w %>%
  mutate(R2_mean = mean(c(event_R2, normal_R2), na.rm=TRUE),
         RMSE_mean = mean(c(event_RMSE, normal_RMSE), na.rm=TRUE),
         coef_mean = mean(c(event_coef, normal_coef), na.rm=TRUE)) %>%
  mutate(R2_mean = coalesce(R2_mean, R2),
         RMSE_mean =  coalesce(event_RMSE, RMSE),
         coef_mean =  coalesce(coef_mean, coef))

spatial_weighted_avg_result <- spatial_table_w %>% 
  select(where(~is.numeric(.x))) %>% 
  summarise(across(!all_of("n") , ~ weighted.mean(.x, n, na.rm = TRUE)))

spatial_weighted_test_result <- spatial_weighted_test_result %>%
  add_row(bind_cols(as_tibble_row(c(Model = 'Everything w.o Event Indicator',
                                    model_response_name = response,
                                    daterange = daterange, lag = TRUE, coordinates = TRUE)),
          as_tibble_row(spatial_weighted_avg_result%>%unlist())))

spatial_weighted_avg_result
```

#### Leave-one-out temporal test
```{r}
# temporal_test <- temporal_validation(curr_model, data, 'Everything w.o Event Indicator',
                                     # response, daterange, weight_max = temporal_weight_max)

# saveRDS(temporal_test, paste0('../../tests/temporal_test/temporal_test_', response, '_', daterange, '_wgt', temporal_weight_max,'.rds'))
temporal_test <- readRDS(paste0('../../tests/temporal_test/temporal_test_', response, '_', daterange, '_wgt', temporal_weight_max,'.rds'))

temporal_test_result <- temporal_test_result %>%
  add_row(temporal_test$result)

temporal_test$result
```

```{r}
temporal_test$plot
```

```{r}
ha_log_full_loeo_plot <- temporal_test$plot
```

# XGB Model performances {.tabset .tabset-pills}
## 80/20 Test Result
```{r}
base_test_result <- test_result %>%
  mutate(transformation = if_else(str_detect(model_response_name, '_log'), 'Log', '')) %>%
  filter(transformation == '') %>%
  mutate(response_base = case_when(str_detect(model_response_name, 'daha') ~ 'Daily Avg From Hourly Avg',
                                   str_detect(model_response_name, 'dmha') ~ 'Daily Max From Hourly Avg',
                                   str_detect(model_response_name, 'ha') ~ 'Hourly Avg')) %>%
  arrange(factor(Model, levels = unique(test_result$Model))) %>%
  select(-transformation, -contains('BT_'))

log_test_result <- test_result %>%
  mutate(transformation = if_else(str_detect(model_response_name, '_log'), 'Log', '')) %>%
  filter(transformation == 'Log') %>%
  mutate(response_base = case_when(str_detect(model_response_name, 'daha') ~ 'Daily Avg From Hourly Avg',
                                   str_detect(model_response_name, 'dmha') ~ 'Daily Max From Hourly Avg',
                                   str_detect(model_response_name, 'ha') ~ 'Hourly Avg'))

base_test_result_unstrat <- test_result_unstrat %>%
  mutate(transformation = if_else(str_detect(model_response_name, '_log'), 'Log', '')) %>%
  filter(transformation == '') %>%
  mutate(response_base = case_when(str_detect(model_response_name, 'daha') ~ 'Daily Avg From Hourly Avg',
                                   str_detect(model_response_name, 'dmha') ~ 'Daily Max From Hourly Avg',
                                   str_detect(model_response_name, 'ha') ~ 'Hourly Avg')) %>%
  arrange(factor(Model, levels = unique(test_result$Model))) %>%
  select(-transformation, -contains('BT_'))

log_test_result_unstrat <- test_result_unstrat %>%
  mutate(transformation = if_else(str_detect(model_response_name, '_log'), 'Log', '')) %>%
  filter(transformation == 'Log') %>%
  mutate(response_base = case_when(str_detect(model_response_name, 'daha') ~ 'Daily Avg From Hourly Avg',
                                   str_detect(model_response_name, 'dmha') ~ 'Daily Max From Hourly Avg',
                                   str_detect(model_response_name, 'ha') ~ 'Hourly Avg'))
```

```{r}
daterange_order <- c('excl_dis', 'full_unstrat')
response_order <- c('Daily Avg From Hourly Avg', 'Daily Max From Hourly Avg', 'Hourly Avg')

# XGB Plot
top <- base_test_result %>%
  filter(Model=='Exclude Event') %>% 
  # rbind(base_test_result_unstrat %>% filter(Model == 'Everything w.o Event Indicator')) %>%
  rbind(base_test_result_unstrat)  %>%
  mutate(response_base = factor(response_base, levels = response_order),
         daterange = factor(daterange, levels = daterange_order)) %>%
  arrange(daterange, response_base) %>%
  mutate(n = c(rep(nrow(daily_daha_excl_dis_test_Xmat), 2),
               nrow(hourly_excl_dis_test_Xmat),
               rep(nrow(daily_daha_full_test_Xmat), 2),
               nrow(hourly_full_test_Xmat))) %>%
  select('response_base', 'daterange', 'n', 'R2', 'RMSE', 'event_R2', 
         'event_RMSE', 'normal_R2', 'normal_RMSE')

bottom <- log_test_result %>%
  filter(Model=='Exclude Event') %>% 
  # rbind(log_test_result_unstrat %>% filter(Model == 'Everything w.o Event Indicator')) %>%
  rbind(log_test_result_unstrat)  %>%
  mutate(response_base = factor(response_base, levels = response_order),
         daterange = factor(daterange, levels = daterange_order)) %>%
  arrange(daterange, response_base) %>%
  mutate(n = c(rep(nrow(daily_daha_excl_dis_test_Xmat), 2),
               nrow(hourly_excl_dis_test_Xmat),
               rep(nrow(daily_daha_full_test_Xmat), 2),
               nrow(hourly_full_test_Xmat))) %>%
  relocate('n', .after = 'daterange') %>%
  select('response_base', 'daterange', 'n', 'BT_R2', 'BT_RMSE', 'BT_event_R2', 
         'BT_event_RMSE', 'BT_normal_R2', 'BT_normal_RMSE') %>%
  rename_with(~ gsub("BT_", "", .x, fixed = TRUE))

xgb_result_table <- bind_rows(top,bottom) %>%
  relocate('daterange') %>%
  setNames(c("Time period", "Response", "N", rep(c("R2", "RMSE"), 3)))

xgb_result_table %>%
  knitr::kable(format = 'pipe', digits = 4, table.attr = "style='width:100%;'") %>%
    pack_rows(index = c("No Transformation" = 6, "Log Transformation" = 6)) %>%
  pack_rows(index = c("Exclude Event" = 3, "Everything w.o D.I" = 3, 
                      "Exclude Event" = 3, "Everything w.o D.I" = 3)) %>%
  add_header_above(c(' ' = 3, 'Full' = 2, 'Event' = 2, 'Routine' = 2)) %>%
  kable_styling()
```

## Spatial leave-one-out Test Results {.tabset .tabset-pills}
### With spatial lag {.tabset .tabset-pills} 
#### Unweighted {.tabset .tabset-pills} 
##### Full time period metrics

```{r}
base_spatial_test_result <- spatial_test_result %>%
  filter(lag == TRUE, coordinates == TRUE) %>%
  mutate(transformation = if_else(str_detect(model_response_name, '_log'), 'Log', '')) %>%
  filter(transformation == '') %>%
  mutate(response_base = case_when(str_detect(model_response_name, 'daha') ~ 'Daily Avg From Hourly Avg',
                                   str_detect(model_response_name, 'dmha') ~ 'Daily Max From Hourly Avg',
                                   str_detect(model_response_name, 'ha') ~ 'Hourly Avg')) %>%
  arrange(factor(Model, levels = unique(spatial_test_result$Model))) %>%
  mutate(response_base = factor(response_base, levels = response_order),
         daterange = factor(daterange, levels = daterange_order)) %>%
  arrange(daterange, response_base)

log_spatial_test_result <- spatial_test_result %>%
  filter(lag == TRUE, coordinates == TRUE) %>%
  mutate(transformation = if_else(str_detect(model_response_name, '_log'), 'Log', '')) %>%
  filter(transformation == 'Log') %>%
  mutate(response_base = case_when(str_detect(model_response_name, 'daha') ~ 'Daily Avg From Hourly Avg',
                                   str_detect(model_response_name, 'dmha') ~ 'Daily Max From Hourly Avg',
                                   str_detect(model_response_name, 'ha') ~ 'Hourly Avg')) %>%
  arrange(factor(Model, levels = unique(spatial_test_result$Model))) %>%
  mutate(response_base = factor(response_base, levels = response_order),
         daterange = factor(daterange, levels = daterange_order)) %>%
  arrange(daterange, response_base)

top <- base_spatial_test_result %>%
  # filter(Model %in% c('Exclude Event', 'Everything w.o Event Indicator')) %>% 
  select('response_base', 'daterange', 'R2', 'RMSE', 'event_R2', 
         'event_RMSE', 'normal_R2', 'normal_RMSE')

bottom <- log_spatial_test_result %>%
  # filter(Model %in% c('Exclude Event', 'Everything w.o Event Indicator')) %>% 
  select('response_base', 'daterange', 'R2', 'RMSE', 'event_R2', 
         'event_RMSE', 'normal_R2', 'normal_RMSE')

spatial_result_table <- bind_rows(top,bottom) %>%
  relocate('daterange') %>%
  setNames(c("Time period", "Response",  rep(c("R2", "RMSE"), 3)))

spatial_result_table %>%
  knitr::kable(format = 'pipe', digits = 4, table.attr = "style='width:100%;'") %>%
    pack_rows(index = c("No Transformation" = 6, "Log Transformation" = 6)) %>%
  pack_rows(index = c("Exclude Event" = 3, "Everything w.o D.I" = 3, 
                      "Exclude Event" = 3, "Everything w.o D.I" = 3)) %>%
  add_header_above(c(' ' = 2, 'Full' = 2, 'Event' = 2, 'Routine' = 2)) %>%
  kable_styling()
```

##### Average of Average of event and routine metrics

```{r}
base_spatial_test_result <- spatial_test_result %>%
  filter(lag == TRUE, coordinates == TRUE) %>%
  mutate(transformation = if_else(str_detect(model_response_name, '_log'), 'Log', '')) %>%
  filter(transformation == '') %>%
  mutate(response_base = case_when(str_detect(model_response_name, 'daha') ~ 'Daily Avg From Hourly Avg',
                                   str_detect(model_response_name, 'dmha') ~ 'Daily Max From Hourly Avg',
                                   str_detect(model_response_name, 'ha') ~ 'Hourly Avg')) %>%
  arrange(factor(Model, levels = unique(spatial_test_result$Model))) %>%
  mutate(response_base = factor(response_base, levels = response_order),
         daterange = factor(daterange, levels = daterange_order)) %>%
  arrange(daterange, response_base)

log_spatial_test_result <- spatial_test_result %>%
  filter(lag == TRUE, coordinates == TRUE) %>%
  mutate(transformation = if_else(str_detect(model_response_name, '_log'), 'Log', '')) %>%
  filter(transformation == 'Log') %>%
  mutate(response_base = case_when(str_detect(model_response_name, 'daha') ~ 'Daily Avg From Hourly Avg',
                                   str_detect(model_response_name, 'dmha') ~ 'Daily Max From Hourly Avg',
                                   str_detect(model_response_name, 'ha') ~ 'Hourly Avg')) %>%
  arrange(factor(Model, levels = unique(spatial_test_result$Model))) %>%
  mutate(response_base = factor(response_base, levels = response_order),
         daterange = factor(daterange, levels = daterange_order)) %>%
  arrange(daterange, response_base)

top <- base_spatial_test_result %>%
  # filter(Model %in% c('Exclude Event', 'Everything w.o Event Indicator')) %>% 
  select('response_base', 'daterange', 'R2_mean', 'RMSE_mean')

bottom <- log_spatial_test_result %>%
  # filter(Model %in% c('Exclude Event', 'Everything w.o Event Indicator')) %>% 
  select('response_base', 'daterange', 'R2_mean', 'RMSE_mean')

spatial_result_table <- bind_rows(top,bottom) %>%
  relocate('daterange') %>%
  setNames(c("Time period", "Response",  c("R2", "RMSE")))

spatial_result_table_kable <- spatial_result_table %>%
  knitr::kable(format = 'latex', digits = 2) %>%
  pack_rows(index = c("No Transformation" = 6, "Log Transformation" = 6)) %>%
  pack_rows(index = c("Exclude Event" = 3, "Everything w.o D.I" = 3, 
                      "Exclude Event" = 3, "Everything w.o D.I" = 3)) %>%
  add_header_above(c(' ' = 2, 'Full' = 2))

spatial_result_table %>%
  knitr::kable(format = 'pipe', digits = 4, table.attr = "style='width:100%;'") %>%
    pack_rows(index = c("No Transformation" = 6, "Log Transformation" = 6)) %>%
  pack_rows(index = c("Exclude Event" = 3, "Everything w.o D.I" = 3, 
                      "Exclude Event" = 3, "Everything w.o D.I" = 3)) %>%
  add_header_above(c(' ' = 2, 'Full' = 2)) %>%
  kable_styling()
```

#### Weighted {.tabset .tabset-pills}
##### Number of left-out-observations per monitor for exclude event model
```{r}
# Number of observations per monitor per date range
excl_dis_data <- bind_rows(daily_ha_train, daily_ha_test) %>%
  filter(!(year == 2021 & month %in% 10:12))

excl_dis_data %>% 
  group_by(Monitor) %>%
  summarise(n = n())
```

##### Number of left-out-observations per monitor for full time models
```{r}
bind_rows(daily_ha_train, daily_ha_test) %>% 
  group_by(Monitor) %>%
  summarise(n = n())
```

##### Full time period metrics

```{r}
base_spatial_weighted_test_result <- spatial_weighted_test_result %>%
  filter(lag == TRUE, coordinates == TRUE) %>%
  mutate(transformation = if_else(str_detect(model_response_name, '_log'), 'Log', '')) %>%
  filter(transformation == '') %>%
  mutate(response_base = case_when(str_detect(model_response_name, 'daha') ~ 'Daily Avg From Hourly Avg',
                                   str_detect(model_response_name, 'dmha') ~ 'Daily Max From Hourly Avg',
                                   str_detect(model_response_name, 'ha') ~ 'Hourly Avg')) %>%
  arrange(factor(Model, levels = unique(spatial_test_result$Model))) %>%
  mutate(response_base = factor(response_base, levels = response_order),
         daterange = factor(daterange, levels = daterange_order)) %>%
  arrange(daterange, response_base)

log_spatial_weighted_test_result <- spatial_weighted_test_result %>%
  filter(lag == TRUE, coordinates == TRUE) %>%
  mutate(transformation = if_else(str_detect(model_response_name, '_log'), 'Log', '')) %>%
  filter(transformation == 'Log') %>%
  mutate(response_base = case_when(str_detect(model_response_name, 'daha') ~ 'Daily Avg From Hourly Avg',
                                   str_detect(model_response_name, 'dmha') ~ 'Daily Max From Hourly Avg',
                                   str_detect(model_response_name, 'ha') ~ 'Hourly Avg')) %>%
  arrange(factor(Model, levels = unique(spatial_test_result$Model))) %>%
  mutate(response_base = factor(response_base, levels = response_order),
         daterange = factor(daterange, levels = daterange_order)) %>%
  arrange(daterange, response_base)

top <- base_spatial_weighted_test_result %>%
  # filter(Model %in% c('Exclude Event', 'Everything w.o Event Indicator')) %>% 
  select('response_base', 'daterange', 'R2', 'RMSE', 'event_R2', 
         'event_RMSE', 'normal_R2', 'normal_RMSE')

bottom <- log_spatial_weighted_test_result %>%
  # filter(Model %in% c('Exclude Event', 'Everything w.o Event Indicator')) %>% 
  select('response_base', 'daterange', 'R2', 'RMSE', 'event_R2', 
         'event_RMSE', 'normal_R2', 'normal_RMSE')

spatial_weighted_result_table <- bind_rows(top,bottom) %>%
  relocate('daterange') %>%
  setNames(c("Time period", "Response",  rep(c("R2", "RMSE"), 3)))

spatial_weighted_result_table_kable <- spatial_weighted_result_table %>%
  knitr::kable(format = 'latex', digits = 2) %>%
  pack_rows(index = c("No Transformation" = 6, "Log Transformation" = 6)) %>%
  pack_rows(index = c("Exclude Event" = 3, "Everything w.o D.I" = 3, 
                      "Exclude Event" = 3, "Everything w.o D.I" = 3)) %>%
  add_header_above(c(' ' = 2, 'Full' = 2, 'Event' = 2, 'Routine' = 2))

# writeLines(spatial_weighted_result_table_kable, '../figures/spatial_weighted_result_table_kable.tex')

spatial_weighted_result_table %>%
  knitr::kable(format = 'pipe', digits = 4, table.attr = "style='width:100%;'") %>%
    pack_rows(index = c("No Transformation" = 6, "Log Transformation" = 6)) %>%
  pack_rows(index = c("Exclude Event" = 3, "Everything w.o D.I" = 3, 
                      "Exclude Event" = 3, "Everything w.o D.I" = 3)) %>%
  add_header_above(c(' ' = 2, 'Full' = 2, 'Event' = 2, 'Routine' = 2)) %>%
  kable_styling()
```

##### Average of Average of event and routine metrics

```{r}
base_spatial_weighted_test_result <- spatial_weighted_test_result %>%
  filter(lag == TRUE, coordinates == TRUE) %>%
  mutate(transformation = if_else(str_detect(model_response_name, '_log'), 'Log', '')) %>%
  filter(transformation == '') %>%
  mutate(response_base = case_when(str_detect(model_response_name, 'daha') ~ 'Daily Avg From Hourly Avg',
                                   str_detect(model_response_name, 'dmha') ~ 'Daily Max From Hourly Avg',
                                   str_detect(model_response_name, 'ha') ~ 'Hourly Avg')) %>%
  arrange(factor(Model, levels = unique(spatial_weighted_test_result$Model))) %>%
  mutate(response_base = factor(response_base, levels = response_order),
         daterange = factor(daterange, levels = daterange_order)) %>%
  arrange(daterange, response_base)

log_spatial_weighted_test_result <- spatial_weighted_test_result %>%
  filter(lag == TRUE, coordinates == TRUE) %>%
  mutate(transformation = if_else(str_detect(model_response_name, '_log'), 'Log', '')) %>%
  filter(transformation == 'Log') %>%
  mutate(response_base = case_when(str_detect(model_response_name, 'daha') ~ 'Daily Avg From Hourly Avg',
                                   str_detect(model_response_name, 'dmha') ~ 'Daily Max From Hourly Avg',
                                   str_detect(model_response_name, 'ha') ~ 'Hourly Avg')) %>%
  arrange(factor(Model, levels = unique(spatial_weighted_test_result$Model))) %>%
  mutate(response_base = factor(response_base, levels = response_order),
         daterange = factor(daterange, levels = daterange_order)) %>%
  arrange(daterange, response_base)

top <- base_spatial_weighted_test_result %>%
  # filter(Model %in% c('Exclude Event', 'Everything w.o Event Indicator')) %>% 
  select('response_base', 'daterange', 'R2_mean', 'RMSE_mean')

bottom <- log_spatial_weighted_test_result %>%
  # filter(Model %in% c('Exclude Event', 'Everything w.o Event Indicator')) %>% 
  select('response_base', 'daterange', 'R2_mean', 'RMSE_mean')

spatial_weighted_result_table <- bind_rows(top,bottom) %>%
  relocate('daterange') %>%
  setNames(c("Time period", "Response",  c("R2", "RMSE")))

spatial_result_table_kable <- spatial_result_table %>%
  knitr::kable(format = 'latex', digits = 2) %>%
  pack_rows(index = c("No Transformation" = 6, "Log Transformation" = 6)) %>%
  pack_rows(index = c("Exclude Event" = 3, "Everything w.o D.I" = 3, 
                      "Exclude Event" = 3, "Everything w.o D.I" = 3)) %>%
  add_header_above(c(' ' = 2, 'Full' = 2))

spatial_weighted_result_table %>%
  knitr::kable(format = 'pipe', digits = 4, table.attr = "style='width:100%;'") %>%
    pack_rows(index = c("No Transformation" = 6, "Log Transformation" = 6)) %>%
  pack_rows(index = c("Exclude Event" = 3, "Everything w.o D.I" = 3, 
                      "Exclude Event" = 3, "Everything w.o D.I" = 3)) %>%
  add_header_above(c(' ' = 2, 'Full' = 2)) %>%
  kable_styling()
```

## Temporal leave-one-out Test Result
```{r}
small_n_daily <- bind_rows(daily_ha_train, daily_ha_test) %>%
  filter(year == 2021 & month %in% 7:8) %>% nrow()
small_n_hourly <- bind_rows(hourly_train, hourly_test) %>%
  filter(year == 2021 & month %in% 7:8) %>% nrow()


base_temporal_test_result <- temporal_test_result %>%
  mutate(transformation = if_else(str_detect(model_response_name, '_log'), 'Log', '')) %>%
  filter(transformation == '') %>%
  mutate(response_base = case_when(str_detect(model_response_name, 'daha') ~ 'Daily Avg From Hourly Avg',
                                   str_detect(model_response_name, 'dmha') ~ 'Daily Max From Hourly Avg',
                                   str_detect(model_response_name, 'ha') ~ 'Hourly Avg')) %>%
  arrange(factor(Model, levels = unique(temporal_test_result$Model))) %>%
  mutate(response_base = factor(response_base, levels = response_order),
         daterange = factor(daterange, levels = daterange_order)) %>%
  arrange(daterange, response_base)

log_temporal_test_result <- temporal_test_result %>%
  mutate(transformation = if_else(str_detect(model_response_name, '_log'), 'Log', '')) %>%
  filter(transformation == 'Log') %>%
  mutate(response_base = case_when(str_detect(model_response_name, 'daha') ~ 'Daily Avg From Hourly Avg',
                                   str_detect(model_response_name, 'dmha') ~ 'Daily Max From Hourly Avg',
                                   str_detect(model_response_name, 'ha') ~ 'Hourly Avg')) %>%
  arrange(factor(Model, levels = unique(spatial_test_result$Model))) %>%
  mutate(response_base = factor(response_base, levels = response_order),
         daterange = factor(daterange, levels = daterange_order)) %>%
  arrange(daterange, response_base)

top <- base_temporal_test_result %>%
  # filter(Model %in% c('Exclude Event', 'Everything w.o Event Indicator')) %>% 
  mutate(n = rep(c(small_n_daily,
                   small_n_daily,
                   small_n_hourly), times = 2)) %>%
  select('response_base', 'daterange', 'n', 'R2', 'RMSE')

bottom <- log_temporal_test_result %>%
  # filter(Model %in% c('Exclude Event','Everything w.o Event Indicator')) %>% 
  mutate(n = rep(c(small_n_daily,
                   small_n_daily,
                   small_n_hourly), times = 2)) %>%
  relocate('n', .after = 'daterange') %>%
  select('response_base', 'daterange', 'n', 'R2', 'RMSE')

temporal_result_table <- bind_rows(top,bottom) %>%
  relocate('daterange') %>%
  setNames(c("Time period", "Response", "N", c("R2", "RMSE")))


temporal_result_table %>%
  knitr::kable(format = 'pipe', digits = 4, table.attr = "style='width:100%;'") %>%
    pack_rows(index = c("No Transformation" = 6, "Log Transformation" = 6)) %>%
  pack_rows(index = c("Exclude Event" = 3, "Everything w.o D.I" = 3, 
                      "Exclude Event" = 3, "Everything w.o D.I" = 3)) %>%
  add_header_above(c(' ' = 3, 'Small Event' = 2)) %>%
  kable_styling()
```


## Best params
```{r}
bestTune_df <- as_tibble(fit.xgb_daha_full_unstrat$bestTune) %>%
  slice_head(n=0) %>%
  mutate(date= character(),
         response = character())
response_names <- c("daha", "daha_log", "dmha", "dmha_log", "ha","ha_log"  )
dateranges <- c("excl_dis", "full_unstrat")
for (date in dateranges) {
  for (response in response_names) {
    model <- get(paste0('fit.xgb_', response, '_', date))
    bestTune_df <- bestTune_df %>%
      bind_rows(as_tibble(model$bestTune) %>% mutate(date = date, response = response))
  }
}
bestTune_df<-bestTune_df %>% relocate(date, response)
bestTune_df
```



